---
title: Best practices for your confirmatory factor analysis
subtitle: A JASP and lavaan tutorial
author:
  - name: Pablo Rogers
    orcid: 0000-0002-0093-3834
    corresponding: true
    email: pablorogers@ufu.br
    affiliations:
      - Universidade Federal de Uberlândia
keywords:
  - Confirmatory Factor Analysis
  - Structural Equation Modeling
  - Internal Structural Validity
  - JASP
  - lavaan
abstract: |
  Confirmatory factor analysis (CFA) is a fundamental method for evaluating the internal structural validity of measurement instruments. In most CFA applications, the measurement model serves as a means to an end rather than an end in itself. To select the appropriate model, prior validity evidence is crucial, and items are typically assessed on an ordinal scale, which has been used in the applied social sciences. However, textbooks on structural equation modeling (SEM) often overlook this specific case, focusing on applications estimable using maximum likelihood (ML) instead. Unfortunately, several popular commercial SEM software packages lack suitable solutions for handling this ‘typical CFA’, leading to confusion and suboptimal decision-making when conducting CFA in this context. This article conceptually contributes to this ongoing discussion by presenting a set of guidelines for conducting a typical CFA, drawing from recent empirical research. We provide a practical contribution by introducing and developing a tutorial example within the JASP and lavaan software platforms. Supplementary materials such as videos, files, and scripts are freely available.
#plain-language-summary: |
#  Earthquake data for the island of La Palma from the September 2021 # eruption is found ...
# key-points:
#  - CFA is essential for assessing the internal structural validity of measurement instruments.
#  - JASP and lavaan offer robust tools for conducting CFA with ordinal data.
date: 03/13/2024
bibliography: references.bib
citation:
  container-title: Best practices for your CFA
number-sections: true
---

## Introduction

### Reading the lavaan and other packages ###

```{r}
library(lavaan)
library(semTools)
library(semPlot)
library(psych)
library(simsem)
library(MASS)
library(writexl)
library(dynamic)
```

### Environment preparation

```{r}
options(max.print=1000000) # Increasing the number of lines in the output
set.seed(123456)
```
###### Step 1: ETL ######

### References
Article:
  # Rogers, P. (2022). Best Practices for Your Exploratory Factor Analysis: 
  # A Factor Tutorial. Revista de Administração Contemporânea, 26(6).
  # https://doi.org/10.1590/1982-7849rac2022210085.en
 ## Database:
  # Rogers, P. (2021). Data for “Best Practices for Your Exploratory Factor Analysis: 
  # A Factor Tutorial.” RAC-Revista de Administração Contemporânea. Mendeley Data, V2.
  # https://doi.org/10.17632/rdky78bk8r.2

## Reading the database directly from the internet

```{r}
 # Defining the database URL
url_data <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/b58a7054-4978-4239-81db-00f9efd86e44/WHOQOL_Data.dat"
 # Reading data without variable names
df0 <- read.table(url_data)
 # Defining the URL of variable names
url_labels <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/ae2a5c01-2fc2-43e5-b31c-8db8cc5bff8a/WHOQOL_Labels.txt"
 # Reading variable names in the first column
column_labels <- read.table(url_labels)
 # Extracting the first column
column_labels <- column_labels$V1
 # Assigning variable names
names(df0) <- column_labels
```

## Viewing the database
```{r}
head(df0)
names(df0)
str(df0)
describe(df0)
```

#### Population Model ####

## for Dynamic Fit Index (DFI) calculation and a priori Power Analysis (see below)
## Considering the model resulting from the meta-analysis by Lin and Yao (2022): popModel
 # Lin, L.-C., & Yao, G. (2022). Validation of the factor structure of 
 # the WHOQOL-BREF using meta-analysis of exploration factor analysis and 
 # social network analysis. Psychological Assessment, 34(7), 660–670.
 # https://doi.org/10.1037/pas0001122
## Considering the ordinal nature of the variables
 # for this I consider the thresholds (t) to be equidistant distributed (linearity): 12%,23%,31%,12%
 # it is unlikely that we will have the thersholds available in any previous study;
 # but if a response frequency is available, for each of the classes, on each of the response items,
 # you can find the thresholds from the inverse of normal distribution.
 # Equidistant thresholds considering a 5-point Likert scale (1 to 5) are: -1.2/-.4/.4/1.2
## Using the factor loadings obtained from Lin and Yao (2022) can be questioned, 
 # as most research in the meta-analysis used the criticized default combination
 # (principal components + Kaiser criterion + varimax rotation) - see Rogers (2022).
 # and more... it is evidence coming from an EFA: not a CFA!
## Consider the example (popModel) for teaching purposes
 # Perhaps you will find a robust survey in your country, which made use of WHOQOL, 
 # using CFA for ordinal data (or not!), and have better starting points.

```{r}
popModel <- '
 # define factors and set loadings according to Lin and Yao (2022)
psycho =~ 0.92*Q5_P + 0.81*Q6_P + 0.94*Q7_P + 0.73*Q11_P + 0.75*Q19_P + 0.63*Q26_P
physical =~ 0.84*Q3_F + 0.83*Q4_F + 0.63*Q10_F + 0.71*Q15_F + 0.68*Q16_F 
            + 0.89*Q17_F + 0.61*Q18_F
social =~ 0.94*Q20_S + 0.89*Q21_S + 0.78*Q22_S
environment =~ 0.37*Q8_A + 0.49*Q9_A + 0.79*Q12_A + 0.81*Q13_A + 0.77*Q14_A
               + 0.75*Q23_A + 0.64*Q24_A + 0.73*Q25_A
 # set the maximum acceptable correlation between Q4 and Q3
Q4_F ~~ 0.3*Q3_F
 # set factor variances to 1 and correlations to 0.3
 # according to Lin and Yao (2022), correlations between factors ranged from 0.08 to 0.3
psycho ~~ 1*psycho + 0.3*physical + 0.3*social + 0.3*environment
physical ~~ 1*physical + 0.3*social + 0.3*environment
social ~~ 1*social + 0.3*environment
environment ~~ 1*environment
 # set all item residual variances according defined loadings (1-L^2)
 # except Q3 and Q4 due to the predicted correlation between the two
Q5_P ~~ 0.15*Q5_P
Q6_P ~~ 0.34*Q6_P
Q7_P ~~ 0.12*Q7_P
Q11_P ~~ 0.47*Q11_P
Q19_P ~~ 0.44*Q19_P
Q26_P ~~ 0.6*Q26_P
Q10_F ~~ 0.6*Q10_F
Q15_F ~~ 0.5*Q15_F
Q16_F ~~ 0.54*Q16_F
Q17_F ~~ 0.21*Q17_F
Q18_F ~~ 0.63*Q18_F
Q20_S ~~ 0.12*Q20_S
Q21_S ~~ 0.21*Q21_S
Q22_S ~~ 0.39*Q22_S
Q8_A ~~ 0.86*Q8_A
Q9_A ~~ 0.76*Q9_A
Q12_A ~~ 0.38*Q12_A
Q13_A ~~ 0.34*Q13_A
Q14_A ~~ 0.41*Q14_A
Q23_A ~~ 0.44*Q23_A
Q24_A ~~ 0.59*Q24_A
Q25_A ~~ 0.47*Q25_A
 # set thresholds
Q5_P | -1.2*t1
Q5_P | -0.4*t2
Q5_P | 0.4*t3
Q5_P | 1.2*t4
Q6_P | -1.2*t1
Q6_P | -0.4*t2
Q6_P | 0.4*t3
Q6_P | 1.2*t4
Q7_P | -1.2*t1
Q7_P | -0.4*t2
Q7_P | 0.4*t3
Q7_P | 1.2*t4
Q11_P | -1.2*t1
Q11_P | -0.4*t2
Q11_P | 0.4*t3
Q11_P | 1.2*t4
Q19_P | -1.2*t1
Q19_P | -0.4*t2
Q19_P | 0.4*t3
Q19_P | 1.2*t4
Q26_P | -1.2*t1
Q26_P | -0.4*t2
Q26_P | 0.4*t3
Q26_P | 1.2*t4
Q3_F | -1.2*t1
Q3_F | -0.4*t2
Q3_F | 0.4*t3
Q3_F | 1.2*t4
Q4_F | -1.2*t1
Q4_F | -0.4*t2
Q4_F | 0.4*t3
Q4_F | 1.2*t4
Q10_F | -1.2*t1
Q10_F | -0.4*t2
Q10_F | 0.4*t3
Q10_F | 1.2*t4
Q15_F | -1.2*t1
Q15_F | -0.4*t2
Q15_F | 0.4*t3
Q15_F | 1.2*t4
Q16_F | -1.2*t1
Q16_F | -0.4*t2
Q16_F | 0.4*t3
Q16_F | 1.2*t4
Q17_F | -1.2*t1
Q17_F | -0.4*t2
Q17_F | 0.4*t3
Q17_F | 1.2*t4
Q18_F | -1.2*t1
Q18_F | -0.4*t2
Q18_F | 0.4*t3
Q18_F | 1.2*t4
Q20_S | -1.2*t1
Q20_S | -0.4*t2
Q20_S | 0.4*t3
Q20_S | 1.2*t4
Q21_S | -1.2*t1
Q21_S | -0.4*t2
Q21_S | 0.4*t3
Q21_S | 1.2*t4
Q22_S | -1.2*t1
Q22_S | -0.4*t2
Q22_S | 0.4*t3
Q22_S | 1.2*t4
Q8_A | -1.2*t1
Q8_A | -0.4*t2
Q8_A | 0.4*t3
Q8_A | 1.2*t4
Q9_A | -1.2*t1
Q9_A | -0.4*t2
Q9_A | 0.4*t3
Q9_A | 1.2*t4
Q12_A | -1.2*t1
Q12_A | -0.4*t2
Q12_A | 0.4*t3
Q12_A | 1.2*t4
Q13_A | -1.2*t1
Q13_A | -0.4*t2
Q13_A | 0.4*t3
Q13_A | 1.2*t4
Q14_A | -1.2*t1
Q14_A | -0.4*t2
Q14_A | 0.4*t3
Q14_A | 1.2*t4
Q23_A | -1.2*t1
Q23_A | -0.4*t2
Q23_A | 0.4*t3
Q23_A | 1.2*t4
Q24_A | -1.2*t1
Q24_A | -0.4*t2
Q24_A | 0.4*t3
Q24_A | 1.2*t4
Q25_A | -1.2*t1
Q25_A | -0.4*t2
Q25_A | 0.4*t3
Q25_A | 1.2*t4
'
```

## Dynamic Fit Index
 # Patient! It takes a while due to simulations
dynamic::catHB(popModel, manual = T, n = 1047, plot = T)
 # These cutoff were valid for all subsequent analyses, 
 # as we are assuming that popModel is the correct population model
## We could calculate a DFI for each of the estimated models, 
 # assuming that the model in question would be the correct population model, 
 # as is usually the case (see the referenced McNeish and Wolf papers).
 # But in this case, we would lose comparability, as to date, we do not have DFI for bicfator models.
 # and DFI for hierarchical models are in progress.

###### 4F model ######

### Step 2: Model Specification ###
 # Endogenous variables are correlated by default
 # The items Q3, Q4 and Q26 have already been reversed
 # With items Q4 and Q3 correlated

```{r}
model.SPE1 <- '
psycho =~ Q5_P + Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
'
```

### Step 3: Model Estimation ###
 # With the CFA function
 # With WLSMV/DWLS estimator
 # Fixing the variance to standardize the factor scale
 # All items are ordinal

```{r}
model.EST1 <- cfa(
                  model.SPE1, data = df0, 
                  ordered = TRUE, # all ordinal items
                  std.lv = TRUE, # factors scale
                  estimator = "WLSMV" # estimator
                 )
```
### Step 4: Model Evaluation ###
 # Be careful with Heywood cases: Std.all > 1 or variance < 0 
 # Look at SRMR, RMSEA and CFI to assess overall fit
 # Look at standardized estimates, residual correlations, and MI to assess local fit

```{r} 
summary(model.EST1, fit.measures = TRUE, 
         standardized = TRUE, rsq = TRUE)
```
## Modification indices (MI)
```{r}
modificationindices(model.EST1, sort. = TRUE,
                    minimum.value = 3.84,
                    maximum.number = 20
                    )
```
## Residual correlations
```{r}
residuals(model.EST1, type = "cor") # look for "cor" > 0.1 
residuals(model.EST1, type = "standardized")$cov # look for |z| > 2.4
```
## Reliability coefficient (must be > .7)
 # alpha = Cronbach's ; omega = composite reliability; omega2 = Bentler's
 # omega3 = McDonald's ; avevar = AVE (average variance extracted > 0.5)
```{r}
semTools::reliability(model.EST1)
```
## Preliminary evaluation
 # The overall fit was mediocre, compatible with a Level 2 of misspecification 
 # Local indices indicated:
   # high correlation between physical and psychological
   # low factor loading of Q4 and Q24
   # Warning: we may have problems with Q5 (it is the cause of many high MI)

## Model Diagram 1
```{r}
semPlot::semPaths(object = model.EST1, 
                  what = "std",
                  style = "lisrel",
                  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
                  edge.label.cex = 0.6,  label.cex= 1.3,
                  weighted= FALSE,
                  edge.color = "black",
                  node.width = 0.6,
                  )
```
## Model Diagram 2
```{r}
semPaths(model.EST1, what = "std",
         style = "lisrel",  layout = "circle",
         intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
         edge.label.cex = 1, label.cex= 1,
         weighted= FALSE,
         mar = c(1.5,0.8,3,0.3),
         )
```
###### 4F bifactor ######

### Step 2: Model Specification ###
 # Endogenous variables are not correlated
 # Items Q3, Q4 and Q26 have already been reversed
 # Without considering the correlated items Q4 and Q3 (the G factor affects both items)
 # Without item Q5 due to convergence problems
 # The G factor (QOL) must be first in the model specification

model.SPE2 <- '
QOL =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P +
       Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F +
       Q20_S + Q21_S + Q22_S +
       Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
'
### Step 3: Model Estimation ###
 # With the CFA function
 # With WLSMV/DWLS estimator
 # Fixing the variance to standardize the factor scale
 # All items are ordinal

model.EST2 <- cfa(model.SPE2, data = df0, 
                  orthogonal = TRUE, # uncorrelated factors
                  ordered = TRUE, # all ordinal items
                  std.lv = TRUE, # factors scale
                  estimator = "WLSMV" # estimator
                 )

### Step 4: Model Evaluation ###
 # Be careful with Heywood cases: Std.all > 1 or variance < 0 
 # Look at SRMR, RMSEA and CFI to assess overall fit
 # Look at standardized estimates, residual correlations, and MI to assess local fit

summary(model.EST2, fit.measures = TRUE, 
        standardized = TRUE, rsq = TRUE
        )

## Modification indices (MI)
modificationindices(model.EST2, sort. = TRUE,
                    minimum.value = 3.84,
                    maximum.number = 20
                    )

## Residual correlations
residuals(model.EST2, type = "cor") # look for "cor" > 0.1 
residuals(model.EST2, type = "standardized")$cov # look for |z| > 2.4

## Reliability coefficient (must be > .7)
 # For the G factor it must be:
psych::omegaFromSem(fit = model.EST2)

## Preliminary Evaluation
 # Overall Fit has improved, but also mediocre (Level 2 of DFI)
 # The G factor is statistically relevant and explains a lot of the observed variables
 # The psycho factor loses a lot of importance

## Model Diagram
semPaths(model.EST2, what = "std",
         layout= "tree3",
         bifactor = "QOL",
         intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
         edge.label.cex = 1.1,  label.cex= 1.2, edge.label.position = 0.7,
         mar = c(3,0.8,3,1),
         weighted= FALSE, edge.color = "black"
         )

###### 4F-2nd model ######

### Step 2: Model specification ###
 # There is a suggestive correlation between endogenous variables that may indicate
 # a second order model, mainly between psycho and physics
 # Items Q3, Q4 and Q26 have already been reversed
 # With items Q4 and Q3 correlated
 # Without item Q5 for comparability
 # Fixing psycho variance due to convergence problems

model.SPE3 <- '
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
QOL =~ psycho + physical + social + environment
Q3_F ~~ Q4_F
psycho ~~ 0*psycho
'
### Step 3: Model Estimation ###
 # With CFA function
 # With WLSMV/DWLS estimator
 # All items are ordinal

model.EST3 <- cfa(
                model.SPE3, data = df0, 
                ordered = TRUE, # all items are ordinal
                #std.lv = TRUE, # factor scale
                estimator = "WLSMV" # estimator
                )
 
### Step 4: Model Evaluation ###
 # Be careful with Heywood cases: Std.all > 1 or variance < 0 
 # Look at SRMR, RMSEA and CFI to assess overall fit
 # Look at standardized estimates, residual correlations, and MI to assess local fit

summary(model.EST3, fit.measures = TRUE, 
        standardized = TRUE, rsq = TRUE)

# Modification indices (MI)
modificationindices(model.EST3, sort. = TRUE,
                    minimum.value = 3.84,
                    maximum.number = 20
                    )

# Residual correlations
residuals(model.EST3, type = "cor") # look for "cor" > 0.1 
residuals(model.EST3, type = "standardized")$cov # look for |z| > 2.4

## Reliability coefficient (must be > .7)
 # alpha = Cronbach's ; omega = composite reliability; omega2 = Bentler's
 # omega3 = McDonald's ; avevar = AVE (average variance extracted > .5)
semTools::reliability(model.EST3)
 # Second order factor reliability
semTools::reliabilityL2(model.EST3, "QOL")

## Model Diagram
semPaths(model.EST3, what = "std",
         style = "lisrel",
         intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
         edge.label.cex = 0.8,  label.cex= 1.2, 
         # edge.label.position = 0.75,
         weighted= FALSE, 
         mar = c(4,1,3,1),
         edge.color = "black",
         node.width = 0.8
         )

###### 4F model without item Q5 ######

### Step 2: Model Specification ###
 # Endogenous variables are correlated by default
 # The items Q3, Q4 and Q26 have already been reversed
 # With items Q4 and Q3 correlated
 # Without item Q5 for comparison

model.SPE4 <- '
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
'
### Step 3: Model Estimation ###
 # With the CFA function
 # With WLSMV/DWLS estimator
 # Fixing the variance to standardize the factor scale
 # All items are ordinal

model.EST4 <- cfa(
    model.SPE4, data = df0, 
    ordered = TRUE, # all items are ordinal
    std.lv = TRUE, # factor scale
    estimator = "WLSMV" # estimator
)

### Step 4: Model Evaluation ###
 # Be careful with Heywood cases: Std.all > 1 or variance < 0 
 # Look at the X2/df, CFI, TLI, SRMR and RMSEA to assess overall fit
 # Look at standardized estimates, residual correlations, and MI to assess local fit

summary(model.EST4, fit.measures = TRUE, 
        standardized = TRUE, rsq = TRUE)

# Modification indices (MI)
modificationindices(model.EST4, sort. = TRUE,
                    minimum.value = 3.84,
                    maximum.number = 20
                    )

# Residuals correlations
residuals(model.EST4, type = "cor") # look for "cor" > 0.1 
residuals(model.EST4, type = "standardized")$cov # look for |z| > 2.4

## Reliability coefficient
 # alpha = Cronbach's ; omega = composite reliability; omega2 = Bentler's
 # omega3 = McDonald's ; avevar = AVE (average variance extracted)
semTools::reliability(model.EST4)

## Model Diagram
semPaths(model.EST4, what = "std",
         style = "lisrel",  layout = "circle",
         intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
         edge.label.cex = 1, label.cex= 1,
         weighted= FALSE,
         mar = c(1.5,0.8,3,0.3),
         )

###### Model comparison ######

fit1 <- fitmeasures(object = model.EST1, c("srmr", "rmsea.scaled", "cfi.scaled") 
)

fit2 <- fitmeasures(object = model.EST2, c("srmr", "rmsea.scaled", "cfi.scaled") 
)

fit3 <- fitmeasures(object = model.EST3, c("srmr", "rmsea.scaled", "cfi.scaled") 
)

fit4 <- fitmeasures(object = model.EST4, c("srmr", "rmsea.scaled", "cfi.scaled") 
)

fit1
fit2
fit3
fit4

## General considerations
 # The models have similar Level of misspecification, but model.EST1 it's a little worse.  
 # The bifactor model presents a slightly better fit. 
 # This finding is consistent with the literature, which points out that, 
 # due to the greater number of parameters, this structure generally presents better adjustments.
 # Some authors even argue that it has been used to save some classic scales,
 # which today are not sustainable due to new estimation and factor selection techniques.
 # Bifactor models also have the drawback of assuming that specific factors are not correlated,
 # which generally does not hold up in practice.
 # Therefore, based on empirical evidence, 
 # I would opt for the 4 correlated factor model, without item Q5.

###### Post Hoc Power Analysis (POW) ######

## Considering original 4F model (without item Q5)
 # As we used the lavaan object for the simulations, we preserved the ordinal nature of the items
## With just 1,000 replications to reduce computational time: 
 # Eventually, for a final result, we can consider 5,000 or 10,000 replications
## n = 1047, according to sample size in the database
## Building CI through simulation and accessing power

Rule.of.thumb <- c(srmr=0.06, rmsea.scaled=0.06, cfi.scaled=0.95)

myModelCI <- simsem::sim(1000, n=nrow(df0), 
                         model = model.SPE4,
                         generate = model.EST4,
                         lavaanfun = "cfa",
                         ordered = TRUE,
                         std.lv = TRUE,
                         estimator = "WLSMV",
                         seed = 123456
)

## Summarize the population model used for data generation underlying a result object
summaryPopulation(myModelCI) # Just checking
## Provide summary of an object
summary(myModelCI) # Preliminary
 # Only two thresholds power < 0.8 (thresholds are generally not parameters of interest)
## Provide short summary of an object
summaryShort(myModelCI) # Summary (97.8% of convergence results)  
## Extract fit indices information and get the cutoffs of fit indices given a priori alpha level
getCutoff(myModelCI, 0.05, usedFit = c("srmr", "rmsea.scaled","cfi.scaled" ))
## Find the proportion of fit indices that indicate worse fit than a specified cutoffs.
getPowerFit(myModelCI, cutoff=Rule.of.thumb, usedFit = c("srmr", "rmsea.scaled","cfi.scaled" ))
## Plot sampling distributions of fit indices that visualize power of rejecting datasets underlying misspecified models
plotPowerFit(myModelCI, cutoff = Rule.of.thumb, logistic = FALSE , usedFit = c("srmr", "rmsea.scaled","cfi.scaled" ))
## Plot sampling distributions of fit indices with fit indices cutoffs
plotCutoff(myModelCI, 0.05, usedFit = c("srmr", "rmsea.scaled","cfi.scaled" ))
## Provide summary of parameter estimates and standard error across replications (alpha=.05)
summaryParam(myModelCI, detail = T, digits = 3) 
 # Look for Power (>.8), Coverage (0.91-0.98), Rel Bias (<|0.10|) and Rel SE Bias (<|0.10|) 
 # All parameters of interest outside this range are points of attention
 # thresholds are generally not parameters of interest
## Saving estimated parameters for inspection
myModelparam <- summaryParam(myModelCI, detail = T)
write_xlsx(myModelparam, "C:/myModelparam.xlsx") # Choose the path on your PC