---
title: Best practices for your confirmatory factor analysis
subtitle: A JASP and lavaan tutorial
author:
  - name: Pablo Rogers
    orcid: 0000-0002-0093-3834
    corresponding: true
    email: pablorogers@ufu.br
    affiliations:
      - Universidade Federal de Uberlândia
keywords:
  - Confirmatory Factor Analysis
  - Structural Equation Modeling
  - Internal Structural Validity
  - JASP
  - lavaan
abstract: |
  Confirmatory factor analysis (CFA) is a fundamental method for evaluating the internal structural validity of measurement instruments. In most CFA applications, the measurement model serves as a means to an end rather than an end in itself. To select the appropriate model, prior validity evidence is crucial, and items are typically assessed on an ordinal scale, which has been used in the applied social sciences. However, textbooks on structural equation modeling (SEM) often overlook this specific case, focusing on applications estimable using maximum likelihood (ML) instead. Unfortunately, several popular commercial SEM software packages lack suitable solutions for handling this 'typical CFA', leading to confusion and suboptimal decision-making when conducting CFA in this context. This article conceptually contributes to this ongoing discussion by presenting a set of guidelines for conducting a typical CFA, drawing from recent empirical research. We provide a practical contribution by introducing and developing a tutorial example within the JASP and lavaan software platforms. Supplementary materials such as videos, files, and scripts are freely available.
#plain-language-summary: |
# key-points:
date: 03/13/2024
bibliography: references.bib
citation:
  container-title: Best practices for your CFA
number-sections: true
execute:
  echo: false
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 80
---

```{r}
#| label: setup-index
#| include: false
library(lavaan)
library(semTools)
library(semPlot)
library(psych)
library(simsem)
source("R/formatting-functions.R")
source("R/plot-themes.R")
df0 <- readRDS("data/whoqol_data.rds")
```

## Introduction

This tutorial demonstrates best practices for conducting a confirmatory factor
analysis (CFA) using the WHOQOL-BREF instrument. All analyses use the DWLS
estimator (WLSMV in lavaan) appropriate for ordinal data. See the reference
article [@rogers2022] and database [@rogers2021] for background.

::: {.callout-tip}
## Tutorial Structure

This article presents polished results only. For the full code behind every
result, see the individual analysis notebooks listed in the Notebooks panel, or
browse the [Article Notebook](index-preview.html).
:::

## Data Description

The WHOQOL-BREF dataset (N = 1,047) was obtained from Mendeley Data
[@rogers2021]. It contains 24 items measuring quality of life across four
domains: Psychological, Physical, Social, and Environment. Items Q3, Q4 and Q26
have already been reverse-coded.
[See data loading code](notebooks/01-setup-data.html).

{{< embed notebooks/01-setup-data.qmd#tbl-data-head >}}

{{< embed notebooks/01-setup-data.qmd#tbl-descriptives >}}

## Population Model

For the calculation of the Dynamic Fit Index (DFI) and the *a priori* power
analysis, the model resulting from the meta-analysis by @lin2022, referred to as
*popModel*, is considered. In their study, @lin2022 validated the factor
structure of the WHOQOL-BREF using a meta-analysis of exploratory factor
analyses combined with social network analysis.
[See population model code](notebooks/02-population-model.html).

::: {.callout-important}
The example population model (*popModel*) should be understood as a teaching
illustration. Ideally, one should look for a robust national survey that has
used the WHOQOL with CFA for ordinal data, providing more appropriate starting
values.
:::

Given the ordinal nature of the variables, equidistant thresholds are assumed,
following a linearity assumption, with proportions of 12%, 23%, 31%, and 12%.
For a five‑point Likert scale (1 to 5), typical equidistant thresholds would be
−1.2, −0.4, 0.4, and 1.2.

## 4-Factor CFA Model {#sec-4f}

The main model specifies four correlated latent factors with all items treated
as ordinal. Items Q4 and Q3 are allowed to correlate.
[See full analysis code](notebooks/03-model-4factor.html).

The lavaan model syntax is:

```default
psycho =~ Q5_P + Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
```

### Model Evaluation

```{r}
#| label: summary-4f-index
est_4fa <- readRDS("models/est_4fa.rds")
summary(est_4fa,
  fit.measures = TRUE,
  standardized = TRUE, rsq = TRUE
)
```

{{< embed notebooks/03-model-4factor.qmd#tbl-reliability-4f >}}

### Preliminary Evaluation

-   The overall fit was mediocre, compatible with a Level 2 of misspecification
-   Local indices indicated:
    -   High correlation between physical and psychological
    -   Low factor loading of Q4 and Q24
    -   Warning: we may have problems with Q5 (it is the cause of many high MI)

### Path Diagram

{{< embed notebooks/03-model-4factor.qmd#fig-diagram-4f-tree >}}

## 4-Factor Bifactor Model {#sec-bifactor}

A bifactor model adds a general Quality of Life (QOL) factor loading on all
items alongside the four specific group factors. Factors are orthogonal
(uncorrelated). Item Q5 is excluded due to convergence problems.
[See full analysis code](notebooks/04-model-bifactor.html).

### Model Evaluation

```{r}
#| label: summary-bifactor-index
est_4f_bi <- readRDS("models/est_4f_bi.rds")
summary(est_4f_bi, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

```{r}
#| label: reliability-bifactor-index
psych::omegaFromSem(fit = est_4f_bi)
```

### Preliminary Evaluation

-   Overall Fit has improved, but also mediocre (Level 2 of DFI)
-   The G factor is statistically relevant and explains a lot of the observed
    variables
-   The psycho factor loses a lot of importance

### Path Diagram

{{< embed notebooks/04-model-bifactor.qmd#fig-diagram-bifactor >}}

## Second-Order CFA Model {#sec-2nd}

A second-order (hierarchical) model where a general QOL factor loads on the four
first-order factors. Item Q5 is excluded for comparability. The psycho factor
variance is fixed to zero due to convergence problems.
[See full analysis code](notebooks/05-model-second-order.html).

### Model Evaluation

```{r}
#| label: summary-2or-index
est_4f_2or <- readRDS("models/est_4f_2or.rds")
summary(est_4f_2or, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

{{< embed notebooks/05-model-second-order.qmd#tbl-reliability-2or >}}

{{< embed notebooks/05-model-second-order.qmd#tbl-reliability-2or-l2 >}}

### Path Diagram

{{< embed notebooks/05-model-second-order.qmd#fig-diagram-2or >}}

## 4-Factor Model Without Item Q5 {#sec-4fb}

Based on the preliminary evaluation of the original 4-factor model, item Q5 was
identified as problematic. This section presents the revised model excluding Q5.
[See full analysis code](notebooks/06-model-4factor-revised.html).

### Model Evaluation

```{r}
#| label: summary-4fb-index
est_4fb <- readRDS("models/est_4fb.rds")
summary(est_4fb, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

{{< embed notebooks/06-model-4factor-revised.qmd#tbl-reliability-4fb >}}

### Path Diagram

{{< embed notebooks/06-model-4factor-revised.qmd#fig-diagram-4fb >}}

## Model Comparison {#sec-comparison}

[See comparison code](notebooks/07-model-comparison.html).

{{< embed notebooks/07-model-comparison.qmd#tbl-model-comparison >}}

### General Considerations

The models exhibit a similar level of misspecification, although the original
4-factor model (with all items) performs slightly worse. The bifactor model
shows a marginally better fit, which aligns with findings in the literature.
This is likely due to the greater number of parameters in bifactor structures,
which generally lead to improved model fit. Some authors even argue that
bifactor models have been used to salvage traditional scales that are no longer
sustainable under modern estimation and factor selection techniques.

::: {.callout-important}
A known limitation of bifactor models is the assumption that specific factors
are uncorrelated — an assumption that often does not hold in practice. Based on
empirical evidence, the four correlated factor model excluding item Q5 is
recommended.
:::

## Post Hoc Power Analysis {#sec-power}

Considering the revised four-factor model (without Q5), we preserved the ordinal
nature of the items by using the lavaan object for simulations. To reduce
computational time, we initially ran only 100 replications; however, for final
results, it would be advisable to increase this to 1,000 or even 5,000
replications. The sample size used was 1,047, consistent with the dataset.
Confidence intervals were constructed through simulation, which also allowed us
to assess statistical power.
[See power analysis code](notebooks/08-power-analysis.html).

{{< embed notebooks/08-power-analysis.qmd#tbl-cutoffs >}}

{{< embed notebooks/08-power-analysis.qmd#tbl-power-fit >}}

{{< embed notebooks/08-power-analysis.qmd#fig-power-fit >}}

{{< embed notebooks/08-power-analysis.qmd#fig-cutoff-plot >}}

::: {.callout-tip}
## Interpreting Power Results

Look for Power > .8, Coverage 0.91–0.98, Rel Bias < |0.10| and Rel SE Bias
< |0.10|. All parameters of interest outside this range are points of
attention. Thresholds are generally not parameters of interest.
:::

```{r}
#| label: param-summary-index
my_modelci <- readRDS("models/my_modelci.rds")
summaryParam(my_modelci, detail = TRUE, digits = 3)
```

## References {.unnumbered}

::: {#refs}
:::
