---
title: Best practices for your confirmatory factor analysis
subtitle: A JASP and lavaan tutorial
author:
  - name: Pablo Rogers
    orcid: 0000-0002-0093-3834
    corresponding: true
    email: pablorogers@ufu.br
    affiliations:
      - Universidade Federal de Uberlândia
keywords:
  - Confirmatory Factor Analysis
  - Structural Equation Modeling
  - Internal Structural Validity
  - JASP
  - lavaan
abstract: |
  Confirmatory factor analysis (CFA) is a fundamental method for evaluating the internal structural validity of measurement instruments. In most CFA applications, the measurement model serves as a means to an end rather than an end in itself. To select the appropriate model, prior validity evidence is crucial, and items are typically assessed on an ordinal scale, which has been used in the applied social sciences. However, textbooks on structural equation modeling (SEM) often overlook this specific case, focusing on applications estimable using maximum likelihood (ML) instead. Unfortunately, several popular commercial SEM software packages lack suitable solutions for handling this ‘typical CFA’, leading to confusion and suboptimal decision-making when conducting CFA in this context. This article conceptually contributes to this ongoing discussion by presenting a set of guidelines for conducting a typical CFA, drawing from recent empirical research. We provide a practical contribution by introducing and developing a tutorial example within the JASP and lavaan software platforms. Supplementary materials such as videos, files, and scripts are freely available.
#plain-language-summary: |
# key-points:
date: 03/13/2024
bibliography: references.bib
citation:
  container-title: Best practices for your CFA
number-sections: true
editor: 
  markdown: 
    wrap: 80
---

## Introduction

### Reading the lavaan and other packages

```{r}
library(lavaan)
library(semTools)
library(semPlot)
library(psych)
library(simsem)
library(MASS)
library(dynamic)
```

### Environment preparation

```{r}
options(max.print = 1000000) # Increasing the number of lines in the output
set.seed(123456)
```

### Step 1: ETL

See reference article [@rogers2022] and database [@rogers2021]

### Reading the database from the internet

```{r}
# Defining the database URL
url_data <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/b58a7054-4978-4239-81db-00f9efd86e44/WHOQOL_Data.dat"
# Reading data without variable names
df0 <- read.table(url_data)
# Defining the URL of variable names
url_labels <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/ae2a5c01-2fc2-43e5-b31c-8db8cc5bff8a/WHOQOL_Labels.txt"
# Reading variable names in the first column
column_labels <- read.table(url_labels)
# Extracting the first column
column_labels <- column_labels$V1
# Assigning variable names
names(df0) <- column_labels
```

### Viewing the database

```{r}
head(df0)
names(df0)
str(df0)
describe(df0)
```

### Population Model

For the calculation of the Dynamic Fit Index (DFI) and the *a priori* power
analysis, the model resulting from the meta-analysis by @lin2022, referred to as
*popModel*, is considered. In their study, lin2022 validated the factor
structure of the WHOQOL-BREF using a meta-analysis of exploratory factor
analyses combined with social network analysis [@lin2022].

Given the ordinal nature of the variables, equidistant thresholds are assumed,
following a linearity assumption, with proportions of 12%, 23%, 31%, and 12%. It
is unlikely that thresholds will be available in previous studies; however, if
response frequencies for each category of each item are provided, thresholds can
be estimated using the inverse normal distribution. For a five‑point Likert
scale (1 to 5), typical equidistant thresholds would be −1.2, −0.4, 0.4, and
1.2.

Using the factor loadings obtained by @lin2022 can be questioned, since most
studies included in their meta-analysis relied on the widely criticized default
combination of principal components, Kaiser criterion, and varimax rotation, as
discussed by @rogers2022. Moreover, the evidence comes from exploratory factor
analysis rather than confirmatory factor analysis.

Therefore, the example model (*popModel*) should be understood as a teaching
illustration. Ideally, one should look for a robust national survey that has
used the WHOQOL with CFA for ordinal data (or otherwise), providing more
appropriate starting values.

```{r}
pop_model <- "
 # define factors and set loadings according to Lin and Yao (2022)
psycho =~ 0.92*Q5_P + 0.81*Q6_P + 0.94*Q7_P + 0.73*Q11_P + 0.75*Q19_P + 0.63*Q26_P
physical =~ 0.84*Q3_F + 0.83*Q4_F + 0.63*Q10_F + 0.71*Q15_F + 0.68*Q16_F
            + 0.89*Q17_F + 0.61*Q18_F
social =~ 0.94*Q20_S + 0.89*Q21_S + 0.78*Q22_S
environment =~ 0.37*Q8_A + 0.49*Q9_A + 0.79*Q12_A + 0.81*Q13_A + 0.77*Q14_A
               + 0.75*Q23_A + 0.64*Q24_A + 0.73*Q25_A
 # set the maximum acceptable correlation between Q4 and Q3
Q4_F ~~ 0.3*Q3_F
 # set factor variances to 1 and correlations to 0.3
 # according to Lin and Yao (2022), correlations between factors ranged from 0.08 to 0.3
psycho ~~ 1*psycho + 0.3*physical + 0.3*social + 0.3*environment
physical ~~ 1*physical + 0.3*social + 0.3*environment
social ~~ 1*social + 0.3*environment
environment ~~ 1*environment
 # set all item residual variances according defined loadings (1-L^2)
 # except Q3 and Q4 due to the predicted correlation between the two
Q5_P ~~ 0.15*Q5_P
Q6_P ~~ 0.34*Q6_P
Q7_P ~~ 0.12*Q7_P
Q11_P ~~ 0.47*Q11_P
Q19_P ~~ 0.44*Q19_P
Q26_P ~~ 0.6*Q26_P
Q10_F ~~ 0.6*Q10_F
Q15_F ~~ 0.5*Q15_F
Q16_F ~~ 0.54*Q16_F
Q17_F ~~ 0.21*Q17_F
Q18_F ~~ 0.63*Q18_F
Q20_S ~~ 0.12*Q20_S
Q21_S ~~ 0.21*Q21_S
Q22_S ~~ 0.39*Q22_S
Q8_A ~~ 0.86*Q8_A
Q9_A ~~ 0.76*Q9_A
Q12_A ~~ 0.38*Q12_A
Q13_A ~~ 0.34*Q13_A
Q14_A ~~ 0.41*Q14_A
Q23_A ~~ 0.44*Q23_A
Q24_A ~~ 0.59*Q24_A
Q25_A ~~ 0.47*Q25_A
 # set thresholds
Q5_P | -1.2*t1
Q5_P | -0.4*t2
Q5_P | 0.4*t3
Q5_P | 1.2*t4
Q6_P | -1.2*t1
Q6_P | -0.4*t2
Q6_P | 0.4*t3
Q6_P | 1.2*t4
Q7_P | -1.2*t1
Q7_P | -0.4*t2
Q7_P | 0.4*t3
Q7_P | 1.2*t4
Q11_P | -1.2*t1
Q11_P | -0.4*t2
Q11_P | 0.4*t3
Q11_P | 1.2*t4
Q19_P | -1.2*t1
Q19_P | -0.4*t2
Q19_P | 0.4*t3
Q19_P | 1.2*t4
Q26_P | -1.2*t1
Q26_P | -0.4*t2
Q26_P | 0.4*t3
Q26_P | 1.2*t4
Q3_F | -1.2*t1
Q3_F | -0.4*t2
Q3_F | 0.4*t3
Q3_F | 1.2*t4
Q4_F | -1.2*t1
Q4_F | -0.4*t2
Q4_F | 0.4*t3
Q4_F | 1.2*t4
Q10_F | -1.2*t1
Q10_F | -0.4*t2
Q10_F | 0.4*t3
Q10_F | 1.2*t4
Q15_F | -1.2*t1
Q15_F | -0.4*t2
Q15_F | 0.4*t3
Q15_F | 1.2*t4
Q16_F | -1.2*t1
Q16_F | -0.4*t2
Q16_F | 0.4*t3
Q16_F | 1.2*t4
Q17_F | -1.2*t1
Q17_F | -0.4*t2
Q17_F | 0.4*t3
Q17_F | 1.2*t4
Q18_F | -1.2*t1
Q18_F | -0.4*t2
Q18_F | 0.4*t3
Q18_F | 1.2*t4
Q20_S | -1.2*t1
Q20_S | -0.4*t2
Q20_S | 0.4*t3
Q20_S | 1.2*t4
Q21_S | -1.2*t1
Q21_S | -0.4*t2
Q21_S | 0.4*t3
Q21_S | 1.2*t4
Q22_S | -1.2*t1
Q22_S | -0.4*t2
Q22_S | 0.4*t3
Q22_S | 1.2*t4
Q8_A | -1.2*t1
Q8_A | -0.4*t2
Q8_A | 0.4*t3
Q8_A | 1.2*t4
Q9_A | -1.2*t1
Q9_A | -0.4*t2
Q9_A | 0.4*t3
Q9_A | 1.2*t4
Q12_A | -1.2*t1
Q12_A | -0.4*t2
Q12_A | 0.4*t3
Q12_A | 1.2*t4
Q13_A | -1.2*t1
Q13_A | -0.4*t2
Q13_A | 0.4*t3
Q13_A | 1.2*t4
Q14_A | -1.2*t1
Q14_A | -0.4*t2
Q14_A | 0.4*t3
Q14_A | 1.2*t4
Q23_A | -1.2*t1
Q23_A | -0.4*t2
Q23_A | 0.4*t3
Q23_A | 1.2*t4
Q24_A | -1.2*t1
Q24_A | -0.4*t2
Q24_A | 0.4*t3
Q24_A | 1.2*t4
Q25_A | -1.2*t1
Q25_A | -0.4*t2
Q25_A | 0.4*t3
Q25_A | 1.2*t4
"
```

### Dynamic Fit Index

Patient! It takes a while due to simulations

```{r}
# dynamic::catHB(pop_model, manual = T, n = 1047, plot = T)
```

These cutoff were valid for all subsequent analyses, as we are assuming that
popModel is the correct population model.

We could calculate a DFI for each of the estimated models, assuming that the
model in question would be the correct population model, as is usually the case
(see the referenced McNeish and Wolf papers). But in this case, we would lose
comparability, as to date, we do not have DFI for bicfator models, and DFI for
hierarchical models are in progress.

## 4F model

### Step 2: Model Specification

-   Endogenous variables are correlated by default
-   The items Q3, Q4 and Q26 have already been reversed
-   With items Q4 and Q3 correlated

```{r}
spe_4fa <- "
psycho =~ Q5_P + Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
"
```

### Step 3: Model Estimation

-   With the CFA function
-   With WLSMV/DWLS estimator
-   Fixing the variance to standardize the factor scale
-   All items are ordinal

```{r}
est_4fa <- cfa(
  spe_4fa,
  data = df0,
  ordered = TRUE, # all ordinal items
  std.lv = TRUE, # factors scale
  estimator = "WLSMV" # estimator
)
```

### Step 4: Model Evaluation

-   Be careful with Heywood cases: Std.all \> 1 or variance \< 0
-   Look at SRMR, RMSEA and CFI to assess overall fit
-   Look at standardized estimates, residual correlations, and MI to assess
    local fit

```{r}
summary(est_4fa,
  fit.measures = TRUE,
  standardized = TRUE, rsq = TRUE
)
```

### Modification indices (MI)

```{r}
modificationindices(est_4fa,
  sort. = TRUE,
  minimum.value = 3.84,
  maximum.number = 20
)
```

### Residual correlations

```{r}
residuals(est_4fa, type = "cor") # look for "cor" > 0.1
residuals(est_4fa, type = "standardized")$cov # look for |z| > 2.4
```

### Reliability coefficient (must be \> .7)

-   alpha = Cronbach's ; omega = composite reliability; omega2 = Bentler's
-   omega3 = McDonald's ; avevar = AVE (average variance extracted > 0.5)

```{r}
semTools::reliability(est_4fa)
```

### Preliminary evaluation

-   The overall fit was mediocre, compatible with a Level 2 of misspecification

-   Local indices indicated:

    -   high correlation between physical and psychological
    -   low factor loading of Q4 and Q24
    -   Warning: we may have problems with Q5 (it is the cause of many high MI)

### Model Diagram 1

```{r}
semPlot::semPaths(
  object = est_4fa,
  what = "std",
  style = "lisrel",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
  edge.label.cex = 0.6, label.cex = 1.3,
  weighted = FALSE,
  edge.color = "black",
  node.width = 0.6,
)
```

### Model Diagram 2

```{r}
semPaths(est_4fa,
  what = "std",
  style = "lisrel", layout = "circle",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
  edge.label.cex = 1, label.cex = 1,
  weighted = FALSE,
  mar = c(1.5, 0.8, 3, 0.3),
)
```

## 4F bifactor

### Step 2: Model Specification

-   Endogenous variables are not correlated
-   Items Q3, Q4 and Q26 have already been reversed
-   Without considering the correlated items Q4 and Q3 (the G factor affects
    both items)
-   Without item Q5 due to convergence problems
-   The G factor (QOL) must be first in the model specification

```{r}
spe_4f_bi <- "
QOL =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P + Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F + Q20_S + Q21_S + Q22_S + Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
"
```

### Step 3: Model Estimation

-   With the CFA function
-   With WLSMV/DWLS estimator
-   Fixing the variance to standardize the factor scale
-   All items are ordinal

```{r}
est_4f_bi <- cfa(spe_4f_bi,
  data = df0,
  orthogonal = TRUE, # uncorrelated factors
  ordered = TRUE, # all ordinal items
  std.lv = TRUE, # factors scale
  estimator = "WLSMV" # estimator
)
```

### Step 4: Model Evaluation

-   Be careful with Heywood cases: Std.all > 1 or variance < 0
-   Look at SRMR, RMSEA and CFI to assess overall fit
-   Look at standardized estimates, residual correlations, and MI to assess
    local fit

```{r}
summary(est_4f_bi, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

### Modification indices (MI)

```{r}
modificationindices(est_4f_bi,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
)
```

### Residual correlations

```{r}
residuals(est_4f_bi, type = "cor") # look for "cor" > 0.1
residuals(est_4f_bi, type = "standardized")$cov # look for |z| > 2.4
```

### Reliability coefficient (must be > .7)

-   For the G factor it must be:

```{r}
psych::omegaFromSem(fit = est_4f_bi)
```

### Preliminary Evaluation

-   Overall Fit has improved, but also mediocre (Level 2 of DFI)
-   The G factor is statistically relevant and explains a lot of the observed
    variables
-   The psycho factor loses a lot of importance

### Model Diagram

```{r}
semPaths(est_4f_bi,
  what = "std", layout = "tree3", bifactor = "QOL",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE, edge.label.cex = 1.1, label.cex =
    1.2, edge.label.position = 0.7, mar = c(3, 0.8, 3, 1), weighted = FALSE,
  edge.color = "black"
)
```

## 4F-2nd model

### Step 2: Model specification

-   There is a suggestive correlation between endogenous variables that may
    indicate a second order model, mainly between psycho and physics
-   Items Q3, Q4 and Q26 have already been reversed
-   With items Q4 and Q3 correlated
-   Without item Q5 for comparability
-   Fixing psycho variance due to convergence problems

```{r}
spe_4f_2or <- "
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
QOL =~ psycho + physical + social + environment
Q3_F ~~ Q4_F
psycho ~~ 0*psycho
"
```

### Step 3: Model Estimation

-   With CFA function
-   With WLSMV/DWLS estimator
-   All items are ordinal

```{r}
est_4f_2or <- cfa(
  spe_4f_2or,
  data = df0,
  ordered = TRUE, # all items are ordinal
  # std.lv = TRUE, # factor scale
  estimator = "WLSMV" # estimator
)
```

### Step 4: Model Evaluation

-   Be careful with Heywood cases: Std.all > 1 or variance < 0
-   Look at SRMR, RMSEA and CFI to assess overall fit
-   Look at standardized estimates, residual correlations, and MI to assess
    local fit

```{r}
summary(est_4f_2or, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

### Modification indices (MI)

```{r}
modificationindices(est_4f_2or,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
)
```

### Residual correlations

```{r}
residuals(est_4f_2or, type = "cor") # look for "cor" > 0.1
residuals(est_4f_2or, type = "standardized")$cov # look for |z| > 2.4
```

### Reliability coefficient (must be \> .7)

-   alpha = Cronbach's ; omega = composite reliability; omega2 = Bentler's
-   omega3 = McDonald's ; avevar = AVE (average variance extracted \> .5)

```{r}
semTools::reliability(est_4f_2or) # Second order factor reliability
semTools::reliabilityL2(est_4f_2or, "QOL")
```

### Model Diagram

```{r}
semPaths(est_4f_2or,
  what = "std",
  style = "lisrel",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
  edge.label.cex = 0.8, label.cex = 1.2,
  weighted = FALSE,
  mar = c(4, 1, 3, 1),
  edge.color = "black",
  node.width = 0.8
)
```

## 4F model without item Q5

### Step 2: Model Specification

-   Endogenous variables are correlated by default
-   The items Q3, Q4 and Q26 have already been reversed
-   With items Q4 and Q3 correlated
-   Without item Q5 for comparison

```{r}
spe_4fb <- "
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
"
```

### Step 3: Model Estimation

-   With the CFA function
-   With WLSMV/DWLS estimator
-   Fixing the variance to standardize the factor scale
-   All items are ordinal

```{r}
est_4fb <- cfa(
  spe_4fb,
  data = df0,
  ordered = TRUE, # all items are ordinal
  std.lv = TRUE, # factor scale
  estimator = "WLSMV" # estimator
)
```

### Step 4: Model Evaluation

-   Be careful with Heywood cases: Std.all \> 1 or variance \< 0
-   Look at the X2/df, CFI, TLI, SRMR and RMSEA to assess overall fit
-   Look at standardized estimates, residual correlations, and MI to assess
    local fit

```{r}
summary(est_4fb, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

### Modification indices (MI)

```{r}
modificationindices(est_4fb,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
)
```

### Residuals correlations

```{r}
residuals(est_4fb, type = "cor") # look for "cor" > 0.1
residuals(est_4fb, type = "standardized")$cov # look for |z| > 2.4
```

### Reliability coefficient

-   alpha = Cronbach's ; omega = composite reliability; omega2 = Bentler's'
-   omega3 = McDonald's ; avevar = AVE (average variance extracted)

```{r}
semTools::reliability(est_4fb)
```

### Model Diagram

```{r}
semPaths(est_4fb,
  what = "std",
  style = "lisrel", layout = "circle",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
  edge.label.cex = 1, label.cex = 1,
  weighted = FALSE,
  mar = c(1.5, 0.8, 3, 0.3),
)
```

## Model comparison

```{r}
fit_4fa <- fitmeasures(object = est_4fa, c("srmr", "rmsea.scaled", "cfi.scaled"))

fit_4f_bi <- fitmeasures(object = est_4f_bi, c("srmr", "rmsea.scaled", "cfi.scaled"))

fit_4f_2or <- fitmeasures(object = est_4f_2or, c("srmr", "rmsea.scaled", "cfi.scaled"))

fit_4fb <- fitmeasures(object = est_4fb, c("srmr", "rmsea.scaled", "cfi.scaled"))

fit_4fa
fit_4f_bi
fit_4f_2or
fit_4fb
```

## General considerations

The models exhibit a similar level of misspecification, although model EST1
performs slightly worse. The bifactor model shows a marginally better fit, which
aligns with findings in the literature. This is likely due to the greater number
of parameters in bifactor structures, which generally lead to improved model
fit. Some authors even argue that bifactor models have been used to salvage
traditional scales that are no longer sustainable under modern estimation and
factor selection techniques. However, a known limitation of bifactor models is
the assumption that specific factors are uncorrelated - an assumption that often
does not hold in practice. Therefore, based on empirical evidence, I would favor
the four correlated factor model, excluding item Q5.

## Post Hoc Power Analysis (POW)

Considering the original four-factor model without item Q5, we preserved the
ordinal nature of the items by using the `lavaan` object for the simulations. To
reduce computational time, we initially ran only 100 replications; however,
for final results, it would be advisable to increase this to 1,000 or even
5,000 replications. The sample size used was 1,047, consistent with the
dataset. Confidence intervals were constructed through simulation, which also
allowed us to assess statistical power.

```{r}
rule_of_thumb <- c(srmr = 0.06, rmsea.scaled = 0.06, cfi.scaled = 0.95)
```

```{r}
my_modelci <- simsem::sim(100,
  n = nrow(df0),
  model = spe_4fb,
  generate = est_4fb,
  lavaanfun = "cfa",
  ordered = TRUE,
  std.lv = TRUE,
  estimator = "WLSMV",
  seed = 123456
)
```

-   Summarize the population model used for data generation underlying a result
    object

```{r}
summaryPopulation(my_modelci) # Just checking
```

-   Provide summary of an object

```{r}
summary(my_modelci) # Preliminary
```

Only two thresholds power < 0.8 (thresholds are generally not parameters of
interest)

-   Provide short summary of an object

```{r}
summaryShort(my_modelci) # Summary (97.8% of convergence results)
```

-   Extract fit indices information and get the cutoffs of fit indices given a
    priori alpha level

```{r}
getCutoff(my_modelci, 0.05, usedFit = c("srmr", "rmsea.scaled", "cfi.scaled"))
```

-   Find the proportion of fit indices that indicate worse fit than a specified
    cutoffs.

```{r}
getPowerFit(my_modelci, cutoff = rule_of_thumb, usedFit = c("srmr", "rmsea.scaled", "cfi.scaled"))
```

-   Plot sampling distributions of fit indices that visualize power of rejecting
    datasets underlying misspecified models

```{r}
plotPowerFit(my_modelci, cutoff = rule_of_thumb, logistic = FALSE, usedFit = c("srmr", "rmsea.scaled", "cfi.scaled"))
```

-   Plot sampling distributions of fit indices with fit indices cutoffs

```{r}
plotCutoff(my_modelci, 0.05, usedFit = c("srmr", "rmsea.scaled", "cfi.scaled"))
```

-   Provide summary of parameter estimates and standard error across
    replications (alpha=.05)

```{r}
summaryParam(my_modelci, detail = TRUE, digits = 3)
```

-   Look for Power (>.8), Coverage (0.91-0.98), Rel Bias (<|0.10|) and Rel
    SE Bias (<|0.10|)
-   All parameters of interest outside this range are points of attention
-   thresholds are generally not parameters of interest

## References {.unnumbered}

::: {#refs}
:::
