---
title: Best practices for your confirmatory factor analysis
subtitle: A JASP and lavaan tutorial
author:
  - name: Pablo Rogers
    orcid: 0000-0002-0093-3834
    corresponding: true
    email: pablorogers@ufu.br
    affiliations:
      - Universidade Federal de Uberlândia
keywords:
  - Confirmatory Factor Analysis
  - Structural Equation Modeling
  - Internal Structural Validity
  - JASP
  - lavaan
abstract: |
  Confirmatory factor analysis (CFA) is a fundamental method for evaluating the internal structural validity of measurement instruments. In most CFA applications, the measurement model serves as a means to an end rather than an end in itself. To select the appropriate model, prior validity evidence is crucial, and items are typically assessed on an ordinal scale, which has been used in the applied social sciences. However, textbooks on structural equation modeling (SEM) often overlook this specific case, focusing on applications estimable using maximum likelihood (ML) instead. Unfortunately, several popular commercial SEM software packages lack suitable solutions for handling this 'typical CFA', leading to confusion and suboptimal decision-making when conducting CFA in this context. This article conceptually contributes to this ongoing discussion by presenting a set of guidelines for conducting a typical CFA, drawing from recent empirical research. We provide a practical contribution by introducing and developing a tutorial example within the JASP and lavaan software platforms. Supplementary materials such as videos, files, and scripts are freely available.
#plain-language-summary: |
# key-points:
date: 03/13/2024
bibliography: references.bib
citation:
  container-title: Best practices for your CFA
number-sections: true
editor: 
  markdown: 
    wrap: 80
---

## Introduction

### Loading packages

```{r}
#| label: setup
#| message: false
library(lavaan)
library(semTools)
library(semPlot)
library(psych)
library(simsem)
library(MASS)
library(dynamic)
```

### Loading helper functions

```{r}
#| label: source-helpers
source(here::here("R", "utils-data.R"))
source(here::here("R", "utils-models.R"))
source(here::here("R", "utils-fit.R"))
```

### Environment preparation

```{r}
#| label: env-setup
set.seed(123456)
```

### Step 1: ETL

See reference article [@rogers2022] and database [@rogers2021]

### Reading the database

```{r}
#| label: load-data
df0 <- load_whoqol_data()
```

### Viewing the database

```{r}
#| label: view-data
head(df0)
names(df0)
str(df0)
describe(df0)
```

### Population Model

For the calculation of the Dynamic Fit Index (DFI) and the *a priori* power
analysis, the model resulting from the meta-analysis by @lin2022, referred to as
*popModel*, is considered. In their study, lin2022 validated the factor
structure of the WHOQOL-BREF using a meta-analysis of exploratory factor
analyses combined with social network analysis [@lin2022].

Given the ordinal nature of the variables, equidistant thresholds are assumed,
following a linearity assumption, with proportions of 12%, 23%, 31%, and 12%. It
is unlikely that thresholds will be available in previous studies; however, if
response frequencies for each category of each item are provided, thresholds can
be estimated using the inverse normal distribution. For a five‑point Likert
scale (1 to 5), typical equidistant thresholds would be −1.2, −0.4, 0.4, and
1.2.

Using the factor loadings obtained by @lin2022 can be questioned, since most
studies included in their meta-analysis relied on the widely criticized default
combination of principal components, Kaiser criterion, and varimax rotation, as
discussed by @rogers2022. Moreover, the evidence comes from exploratory factor
analysis rather than confirmatory factor analysis.

Therefore, the example model (*popModel*) should be understood as a teaching
illustration. Ideally, one should look for a robust national survey that has
used the WHOQOL with CFA for ordinal data (or otherwise), providing more
appropriate starting values.

```{r}
#| label: pop-model
pop_model <- specify_pop_model()
```

### Dynamic Fit Index

Patient! It takes a while due to simulations

```{r}
#| label: dfi
# dynamic::catHB(pop_model, manual = TRUE, n = 1047, plot = TRUE)
```

These cutoff were valid for all subsequent analyses, as we are assuming that
popModel is the correct population model.

We could calculate a DFI for each of the estimated models, assuming that the
model in question would be the correct population model, as is usually the case
(see the referenced McNeish and Wolf papers). But in this case, we would lose
comparability, as to date, we do not have DFI for bicfator models, and DFI for
hierarchical models are in progress.

## 4F model

### Step 2: Model Specification

-   Endogenous variables are correlated by default
-   The items Q3, Q4 and Q26 have already been reversed
-   With items Q4 and Q3 correlated

```{r}
#| label: spe-4fa
spe_4fa <- specify_4f_model()
```

### Step 3: Model Estimation

-   With the CFA function
-   With WLSMV/DWLS estimator
-   Fixing the variance to standardize the factor scale
-   All items are ordinal

```{r}
#| label: est-4fa
est_4fa <- cfa(
  spe_4fa,
  data = df0,
  ordered = TRUE,  # all ordinal items
  std.lv = TRUE,   # factor scale standardization
  estimator = "WLSMV"
)
```

### Step 4: Model Evaluation

-   Be careful with Heywood cases: Std.all \> 1 or variance \< 0
-   Look at SRMR, RMSEA and CFI to assess overall fit
-   Look at standardized estimates, residual correlations, and MI to assess
    local fit

```{r}
#| label: eval-4fa
summary(est_4fa,
  fit.measures = TRUE,
  standardized = TRUE, rsq = TRUE
)
```

### Modification indices (MI)

```{r}
#| label: mi-4fa
modificationindices(est_4fa,
  sort. = TRUE,
  minimum.value = 3.84,
  maximum.number = 20
)
```

### Residual correlations

```{r}
#| label: resid-4fa
residuals(est_4fa, type = "cor") # look for "cor" > 0.1
residuals(est_4fa, type = "standardized")$cov # look for |z| > 2.4
```

### Reliability coefficient (must be \> .7)

-   alpha = Cronbach's ; omega = composite reliability; omega2 = Bentler's
-   omega3 = McDonald's ; avevar = AVE (average variance extracted > 0.5)

```{r}
#| label: rel-4fa
semTools::reliability(est_4fa)
```

### Preliminary evaluation

-   The overall fit was mediocre, compatible with a Level 2 of misspecification

-   Local indices indicated:

    -   high correlation between physical and psychological
    -   low factor loading of Q4 and Q24
    -   Warning: we may have problems with Q5 (it is the cause of many high MI)

### Model Diagram 1

```{r}
#| label: fig-4fa-lisrel
#| fig-cap: "4F model — LISREL-style path diagram"
semPlot::semPaths(
  object = est_4fa,
  what = "std",
  style = "lisrel",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
  edge.label.cex = 0.6, label.cex = 1.3,
  weighted = FALSE,
  edge.color = "black",
  node.width = 0.6
)
```

### Model Diagram 2

```{r}
#| label: fig-4fa-circle
#| fig-cap: "4F model — circular layout"
semPaths(est_4fa,
  what = "std",
  style = "lisrel", layout = "circle",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
  edge.label.cex = 1, label.cex = 1,
  weighted = FALSE,
  mar = c(1.5, 0.8, 3, 0.3)
)
```

## 4F bifactor

### Step 2: Model Specification

-   Endogenous variables are not correlated
-   Items Q3, Q4 and Q26 have already been reversed
-   Without considering the correlated items Q4 and Q3 (the G factor affects
    both items)
-   Without item Q5 due to convergence problems
-   The G factor (QOL) must be first in the model specification

```{r}
#| label: spe-bifactor
spe_4f_bi <- specify_bifactor_model()
```

### Step 3: Model Estimation

-   With the CFA function
-   With WLSMV/DWLS estimator
-   Fixing the variance to standardize the factor scale
-   All items are ordinal

```{r}
#| label: est-bifactor
est_4f_bi <- cfa(spe_4f_bi,
  data = df0,
  orthogonal = TRUE, # uncorrelated factors
  ordered = TRUE,    # all ordinal items
  std.lv = TRUE,     # factor scale standardization
  estimator = "WLSMV"
)
```

### Step 4: Model Evaluation

-   Be careful with Heywood cases: Std.all > 1 or variance < 0
-   Look at SRMR, RMSEA and CFI to assess overall fit
-   Look at standardized estimates, residual correlations, and MI to assess
    local fit

```{r}
#| label: eval-bifactor
summary(est_4f_bi, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

### Modification indices (MI)

```{r}
#| label: mi-bifactor
modificationindices(est_4f_bi,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
)
```

### Residual correlations

```{r}
#| label: resid-bifactor
residuals(est_4f_bi, type = "cor") # look for "cor" > 0.1
residuals(est_4f_bi, type = "standardized")$cov # look for |z| > 2.4
```

### Reliability coefficient (must be > .7)

-   For the G factor it must be:

```{r}
#| label: rel-bifactor
psych::omegaFromSem(fit = est_4f_bi)
```

### Preliminary Evaluation

-   Overall Fit has improved, but also mediocre (Level 2 of DFI)
-   The G factor is statistically relevant and explains a lot of the observed
    variables
-   The psycho factor loses a lot of importance

### Model Diagram

```{r}
#| label: fig-bifactor
#| fig-cap: "Bifactor model diagram"
semPaths(est_4f_bi,
  what = "std", layout = "tree3", bifactor = "QOL",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
  edge.label.cex = 1.1, label.cex = 1.2, edge.label.position = 0.7,
  mar = c(3, 0.8, 3, 1), weighted = FALSE,
  edge.color = "black"
)
```

## 4F-2nd model

### Step 2: Model specification

-   There is a suggestive correlation between endogenous variables that may
    indicate a second order model, mainly between psycho and physics
-   Items Q3, Q4 and Q26 have already been reversed
-   With items Q4 and Q3 correlated
-   Without item Q5 for comparability
-   Fixing psycho variance due to convergence problems

```{r}
#| label: spe-2nd
spe_4f_2or <- specify_second_order_model()
```

### Step 3: Model Estimation

-   With CFA function
-   With WLSMV/DWLS estimator
-   All items are ordinal

```{r}
#| label: est-2nd
est_4f_2or <- cfa(
  spe_4f_2or,
  data = df0,
  ordered = TRUE,    # all items are ordinal
  # std.lv = TRUE,   # not used — psycho variance fixed to 0 in model spec
  estimator = "WLSMV"
)
```

### Step 4: Model Evaluation

-   Be careful with Heywood cases: Std.all > 1 or variance < 0
-   Look at SRMR, RMSEA and CFI to assess overall fit
-   Look at standardized estimates, residual correlations, and MI to assess
    local fit

```{r}
#| label: eval-2nd
summary(est_4f_2or, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

### Modification indices (MI)

```{r}
#| label: mi-2nd
modificationindices(est_4f_2or,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
)
```

### Residual correlations

```{r}
#| label: resid-2nd
residuals(est_4f_2or, type = "cor") # look for "cor" > 0.1
residuals(est_4f_2or, type = "standardized")$cov # look for |z| > 2.4
```

### Reliability coefficient (must be \> .7)

-   alpha = Cronbach's ; omega = composite reliability; omega2 = Bentler's
-   omega3 = McDonald's ; avevar = AVE (average variance extracted \> .5)

```{r}
#| label: rel-2nd
semTools::reliability(est_4f_2or) # Second order factor reliability
semTools::reliabilityL2(est_4f_2or, "QOL")
```

### Model Diagram

```{r}
#| label: fig-2nd
#| fig-cap: "Second-order model diagram"
semPaths(est_4f_2or,
  what = "std",
  style = "lisrel",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
  edge.label.cex = 0.8, label.cex = 1.2,
  weighted = FALSE,
  mar = c(4, 1, 3, 1),
  edge.color = "black",
  node.width = 0.8
)
```

## 4F model without item Q5

### Step 2: Model Specification

-   Endogenous variables are correlated by default
-   The items Q3, Q4 and Q26 have already been reversed
-   With items Q4 and Q3 correlated
-   Without item Q5 for comparison

```{r}
#| label: spe-4fb
spe_4fb <- specify_4f_no_q5_model()
```

### Step 3: Model Estimation

-   With the CFA function
-   With WLSMV/DWLS estimator
-   Fixing the variance to standardize the factor scale
-   All items are ordinal

```{r}
#| label: est-4fb
est_4fb <- cfa(
  spe_4fb,
  data = df0,
  ordered = TRUE,  # all items are ordinal
  std.lv = TRUE,   # factor scale standardization
  estimator = "WLSMV"
)
```

### Step 4: Model Evaluation

-   Be careful with Heywood cases: Std.all \> 1 or variance \< 0
-   Look at the X2/df, CFI, TLI, SRMR and RMSEA to assess overall fit
-   Look at standardized estimates, residual correlations, and MI to assess
    local fit

```{r}
#| label: eval-4fb
summary(est_4fb, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

### Modification indices (MI)

```{r}
#| label: mi-4fb
modificationindices(est_4fb,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
)
```

### Residuals correlations

```{r}
#| label: resid-4fb
residuals(est_4fb, type = "cor") # look for "cor" > 0.1
residuals(est_4fb, type = "standardized")$cov # look for |z| > 2.4
```

### Reliability coefficient

-   alpha = Cronbach's ; omega = composite reliability; omega2 = Bentler's'
-   omega3 = McDonald's ; avevar = AVE (average variance extracted)

```{r}
#| label: rel-4fb
semTools::reliability(est_4fb)
```

### Model Diagram

```{r}
#| label: fig-4fb
#| fig-cap: "4F model without Q5 — circular layout"
semPaths(est_4fb,
  what = "std",
  style = "lisrel", layout = "circle",
  intercepts = FALSE, thresholds = FALSE, residuals = FALSE,
  edge.label.cex = 1, label.cex = 1,
  weighted = FALSE,
  mar = c(1.5, 0.8, 3, 0.3)
)
```

## Model comparison

```{r}
#| label: tbl-fit-comparison
#| tbl-cap: "Fit indices comparison across models"
fit_comparison <- compare_fit(
  `4F (with Q5)`    = est_4fa,
  `4F Bifactor`     = est_4f_bi,
  `4F Second-order` = est_4f_2or,
  `4F (no Q5)`      = est_4fb
)
knitr::kable(fit_comparison, digits = 4)
```

## General considerations

The models exhibit a similar level of misspecification, although model EST1
performs slightly worse. The bifactor model shows a marginally better fit, which
aligns with findings in the literature. This is likely due to the greater number
of parameters in bifactor structures, which generally lead to improved model
fit. Some authors even argue that bifactor models have been used to salvage
traditional scales that are no longer sustainable under modern estimation and
factor selection techniques. However, a known limitation of bifactor models is
the assumption that specific factors are uncorrelated - an assumption that often
does not hold in practice. Therefore, based on empirical evidence, I would favor
the four correlated factor model, excluding item Q5.

## Post Hoc Power Analysis (POW)

Considering the original four-factor model without item Q5, we preserved the
ordinal nature of the items by using the `lavaan` object for the simulations. To
reduce computational time, we initially ran only 100 replications; however,
for final results, it would be advisable to increase this to 1,000 or even
5,000 replications. The sample size used was 1,047, consistent with the
dataset. Confidence intervals were constructed through simulation, which also
allowed us to assess statistical power.

```{r}
#| label: power-cutoffs
rule_of_thumb <- c(srmr = 0.06, rmsea.scaled = 0.06, cfi.scaled = 0.95)
```

```{r}
#| label: power-sim
my_modelci <- simsem::sim(100,
  n = nrow(df0),
  model = spe_4fb,
  generate = est_4fb,
  lavaanfun = "cfa",
  ordered = TRUE,
  std.lv = TRUE,
  estimator = "WLSMV",
  seed = 123456
)
```

-   Summarize the population model used for data generation underlying a result
    object

```{r}
#| label: power-pop
summaryPopulation(my_modelci) # Just checking
```

-   Provide summary of an object

```{r}
#| label: power-summary
summary(my_modelci) # Preliminary
```

Only two thresholds power < 0.8 (thresholds are generally not parameters of
interest)

-   Provide short summary of an object

```{r}
#| label: power-short
summaryShort(my_modelci) # Summary (97.8% of convergence results)
```

-   Extract fit indices information and get the cutoffs of fit indices given a
    priori alpha level

```{r}
#| label: power-cutoff-extract
getCutoff(my_modelci, 0.05, usedFit = c("srmr", "rmsea.scaled", "cfi.scaled"))
```

-   Find the proportion of fit indices that indicate worse fit than a specified
    cutoffs.

```{r}
#| label: power-fit
getPowerFit(my_modelci,
  cutoff = rule_of_thumb,
  usedFit = c("srmr", "rmsea.scaled", "cfi.scaled")
)
```

-   Plot sampling distributions of fit indices that visualize power of rejecting
    datasets underlying misspecified models

```{r}
#| label: fig-power-fit
#| fig-cap: "Sampling distributions of fit indices — power analysis"
plotPowerFit(my_modelci,
  cutoff = rule_of_thumb, logistic = FALSE,
  usedFit = c("srmr", "rmsea.scaled", "cfi.scaled")
)
```

-   Plot sampling distributions of fit indices with fit indices cutoffs

```{r}
#| label: fig-power-cutoff
#| fig-cap: "Sampling distributions of fit indices with cutoffs"
plotCutoff(my_modelci, 0.05,
  usedFit = c("srmr", "rmsea.scaled", "cfi.scaled")
)
```

-   Provide summary of parameter estimates and standard error across
    replications (alpha=.05)

```{r}
#| label: power-params
summaryParam(my_modelci, detail = TRUE, digits = 3)
```

-   Look for Power (>.8), Coverage (0.91-0.98), Rel Bias (<|0.10|) and Rel
    SE Bias (<|0.10|)
-   All parameters of interest outside this range are points of attention
-   thresholds are generally not parameters of interest

## Session Information {.unnumbered}

```{r}
#| label: session-info
sessionInfo()
```

## References {.unnumbered}

::: {#refs}
:::
