---
title: Best practices for your confirmatory factor analysis
subtitle: A JASP and lavaan tutorial
author:
  - name: Pablo Rogers
    orcid: 0000-0002-0093-3834
    corresponding: true
    email: pablorogers@ufu.br
    affiliations:
      - Universidade Federal de Uberlândia
keywords:
  - Confirmatory Factor Analysis
  - Structural Equation Modeling
  - Internal Structural Validity
  - JASP
  - lavaan
abstract: |
  Confirmatory factor analysis (CFA) is a fundamental method for evaluating the internal structural validity of measurement instruments. In most CFA applications, the measurement model serves as a means to an end rather than an end in itself. To select the appropriate model, prior validity evidence is crucial, and items are typically assessed on an ordinal scale, which has been used in the applied social sciences. However, textbooks on structural equation modeling (SEM) often overlook this specific case, focusing on applications estimable using maximum likelihood (ML) instead. Unfortunately, several popular commercial SEM software packages lack suitable solutions for handling this 'typical CFA', leading to confusion and suboptimal decision-making when conducting CFA in this context. This article conceptually contributes to this ongoing discussion by presenting a set of guidelines for conducting a typical CFA, drawing from recent empirical research. We provide a practical contribution by introducing and developing a tutorial example within the JASP and lavaan software platforms. Supplementary materials such as videos, files, and scripts are freely available.
date: 03/13/2024
bibliography: references.bib
citation:
  container-title: Best practices for your CFA
number-sections: true
execute:
  echo: true       # Show all code (this is a tutorial)
  warning: false
  message: false
  cache: true      # Enable caching for faster rendering
editor: 
  markdown: 
    wrap: 80
---

## Introduction

Confirmatory factor analysis (CFA) is a fundamental method for evaluating the internal structural validity of measurement instruments. This tutorial demonstrates best practices for conducting CFA using the WHOQOL-BREF instrument, a widely used quality of life assessment tool.

All analyses in this tutorial use the DWLS estimator (called WLSMV in lavaan), which is appropriate for ordinal data. This is the "typical CFA" scenario in applied social sciences, where items are measured on Likert scales rather than continuous variables.

::: {.callout-important}
## Tutorial Objective

This tutorial is designed to be **fully self-contained**. All code, explanations, and results are presented in this single document, making it easy to follow the entire CFA workflow from data loading through model comparison and power analysis.
:::

The WHOQOL-BREF dataset (N = 1,047) was obtained from Mendeley Data [@rogers2021] and supports the methodological framework described in @rogers2022. The instrument contains 24 items measuring quality of life across four domains: Psychological (6 items), Physical (7 items), Social (3 items), and Environment (8 items).

## Setup

First, we load all required R packages for the analysis:

```{r}
#| label: load-packages
library(lavaan)     # CFA estimation and model fitting
library(semTools)   # Reliability and additional SEM tools
library(semPlot)    # Path diagrams
library(psych)      # Descriptive statistics and omega coefficients
library(simsem)     # Power analysis via simulation
library(MASS)       # General statistical functions
library(dynamic)    # Dynamic Fit Index computation
```

Set global options and seed for reproducibility:

```{r}
#| label: set-options
options(max.print = 1000000)  # Allow printing large outputs
set.seed(123456)              # Ensure reproducible results for simulations
```

## Step 1: ETL — Extract, Transform, Load

The WHOQOL-BREF data are loaded directly from Mendeley Data [@rogers2021]. Items Q3, Q4, and Q26 have already been reverse-coded in the source data.

### Loading the Data

```{r}
#| label: load-data
# Define the database URL
url_data <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/b58a7054-4978-4239-81db-00f9efd86e44/WHOQOL_Data.dat"

# Read data without variable names
df0 <- read.table(url_data)

# Define the URL of variable names
url_labels <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/ae2a5c01-2fc2-43e5-b31c-8db8cc5bff8a/WHOQOL_Labels.txt"

# Read variable names from the first column
column_labels <- read.table(url_labels)
column_labels <- column_labels$V1

# Assign variable names to the dataset
names(df0) <- column_labels
```

### Exploring the Data

```{r}
#| label: tbl-data-head
#| tbl-cap: "First rows of the WHOQOL-BREF dataset"
knitr::kable(head(df0))
```

```{r}
#| label: view-variable-names
# Display all variable names
names(df0)
```

```{r}
#| label: view-data-structure
# Display data structure
str(df0)
```

```{r}
#| label: tbl-descriptives
#| tbl-cap: "Descriptive statistics for all WHOQOL-BREF items"
knitr::kable(describe(df0), digits = 2)
```

::: {.callout-note}
## Data Overview

The dataset contains N = `r nrow(df0)` observations on `r ncol(df0)` variables. All items are measured on a 5-point Likert scale (1-5), representing ordinal data. Items Q3, Q4, and Q26 have been reverse-coded so that higher values consistently indicate better quality of life.
:::

## Population Model

For the calculation of the Dynamic Fit Index (DFI) and *a priori* power analysis, we use a population model based on the meta-analysis by @lin2022. In their study, Lin and Yao validated the factor structure of the WHOQOL-BREF using a meta-analysis of exploratory factor analyses combined with social network analysis.

::: {.callout-important}
## Important Caveat

The example population model (*popModel*) should be understood as a **teaching illustration**. Ideally, one should look for a robust national survey that has used the WHOQOL-BREF with CFA for ordinal data, providing more appropriate starting values. The loadings from @lin2022 come primarily from exploratory factor analysis studies using principal components with varimax rotation—a method that has known limitations [@rogers2022].
:::

### Defining the Population Model

Given the ordinal nature of the variables, we assume equidistant thresholds following a linearity assumption, with proportions of 12%, 23%, 31%, and 12%. For a five-point Likert scale (1 to 5), typical equidistant thresholds would be −1.2, −0.4, 0.4, and 1.2. While it is unlikely that thresholds will be available in previous studies, if response frequencies for each category are provided, thresholds can be estimated using the inverse normal distribution.

```{r}
#| label: define-pop-model
pop_model <- "
 # Define factors and set loadings according to Lin and Yao (2022)
psycho =~ 0.92*Q5_P + 0.81*Q6_P + 0.94*Q7_P + 0.73*Q11_P + 0.75*Q19_P + 0.63*Q26_P
physical =~ 0.84*Q3_F + 0.83*Q4_F + 0.63*Q10_F + 0.71*Q15_F + 0.68*Q16_F
            + 0.89*Q17_F + 0.61*Q18_F
social =~ 0.94*Q20_S + 0.89*Q21_S + 0.78*Q22_S
environment =~ 0.37*Q8_A + 0.49*Q9_A + 0.79*Q12_A + 0.81*Q13_A + 0.77*Q14_A
               + 0.75*Q23_A + 0.64*Q24_A + 0.73*Q25_A
 # Set the maximum acceptable correlation between Q4 and Q3
Q4_F ~~ 0.3*Q3_F
 # Set factor variances to 1 and correlations to 0.3
 # According to Lin and Yao (2022), correlations between factors ranged from 0.08 to 0.3
psycho ~~ 1*psycho + 0.3*physical + 0.3*social + 0.3*environment
physical ~~ 1*physical + 0.3*social + 0.3*environment
social ~~ 1*social + 0.3*environment
environment ~~ 1*environment
 # Set all item residual variances according to defined loadings (1-L^2)
 # Except Q3 and Q4 due to the predicted correlation between the two
Q5_P ~~ 0.15*Q5_P
Q6_P ~~ 0.34*Q6_P
Q7_P ~~ 0.12*Q7_P
Q11_P ~~ 0.47*Q11_P
Q19_P ~~ 0.44*Q19_P
Q26_P ~~ 0.6*Q26_P
Q10_F ~~ 0.6*Q10_F
Q15_F ~~ 0.5*Q15_F
Q16_F ~~ 0.54*Q16_F
Q17_F ~~ 0.21*Q17_F
Q18_F ~~ 0.63*Q18_F
Q20_S ~~ 0.12*Q20_S
Q21_S ~~ 0.21*Q21_S
Q22_S ~~ 0.39*Q22_S
Q8_A ~~ 0.86*Q8_A
Q9_A ~~ 0.76*Q9_A
Q12_A ~~ 0.38*Q12_A
Q13_A ~~ 0.34*Q13_A
Q14_A ~~ 0.41*Q14_A
Q23_A ~~ 0.44*Q23_A
Q24_A ~~ 0.59*Q24_A
Q25_A ~~ 0.47*Q25_A
 # Set equidistant thresholds for all items
Q5_P | -1.2*t1
Q5_P | -0.4*t2
Q5_P | 0.4*t3
Q5_P | 1.2*t4
Q6_P | -1.2*t1
Q6_P | -0.4*t2
Q6_P | 0.4*t3
Q6_P | 1.2*t4
Q7_P | -1.2*t1
Q7_P | -0.4*t2
Q7_P | 0.4*t3
Q7_P | 1.2*t4
Q11_P | -1.2*t1
Q11_P | -0.4*t2
Q11_P | 0.4*t3
Q11_P | 1.2*t4
Q19_P | -1.2*t1
Q19_P | -0.4*t2
Q19_P | 0.4*t3
Q19_P | 1.2*t4
Q26_P | -1.2*t1
Q26_P | -0.4*t2
Q26_P | 0.4*t3
Q26_P | 1.2*t4
Q3_F | -1.2*t1
Q3_F | -0.4*t2
Q3_F | 0.4*t3
Q3_F | 1.2*t4
Q4_F | -1.2*t1
Q4_F | -0.4*t2
Q4_F | 0.4*t3
Q4_F | 1.2*t4
Q10_F | -1.2*t1
Q10_F | -0.4*t2
Q10_F | 0.4*t3
Q10_F | 1.2*t4
Q15_F | -1.2*t1
Q15_F | -0.4*t2
Q15_F | 0.4*t3
Q15_F | 1.2*t4
Q16_F | -1.2*t1
Q16_F | -0.4*t2
Q16_F | 0.4*t3
Q16_F | 1.2*t4
Q17_F | -1.2*t1
Q17_F | -0.4*t2
Q17_F | 0.4*t3
Q17_F | 1.2*t4
Q18_F | -1.2*t1
Q18_F | -0.4*t2
Q18_F | 0.4*t3
Q18_F | 1.2*t4
Q20_S | -1.2*t1
Q20_S | -0.4*t2
Q20_S | 0.4*t3
Q20_S | 1.2*t4
Q21_S | -1.2*t1
Q21_S | -0.4*t2
Q21_S | 0.4*t3
Q21_S | 1.2*t4
Q22_S | -1.2*t1
Q22_S | -0.4*t2
Q22_S | 0.4*t3
Q22_S | 1.2*t4
Q8_A | -1.2*t1
Q8_A | -0.4*t2
Q8_A | 0.4*t3
Q8_A | 1.2*t4
Q9_A | -1.2*t1
Q9_A | -0.4*t2
Q9_A | 0.4*t3
Q9_A | 1.2*t4
Q12_A | -1.2*t1
Q12_A | -0.4*t2
Q12_A | 0.4*t3
Q12_A | 1.2*t4
Q13_A | -1.2*t1
Q13_A | -0.4*t2
Q13_A | 0.4*t3
Q13_A | 1.2*t4
Q14_A | -1.2*t1
Q14_A | -0.4*t2
Q14_A | 0.4*t3
Q14_A | 1.2*t4
Q23_A | -1.2*t1
Q23_A | -0.4*t2
Q23_A | 0.4*t3
Q23_A | 1.2*t4
Q24_A | -1.2*t1
Q24_A | -0.4*t2
Q24_A | 0.4*t3
Q24_A | 1.2*t4
Q25_A | -1.2*t1
Q25_A | -0.4*t2
Q25_A | 0.4*t3
Q25_A | 1.2*t4
"
```

### Dynamic Fit Index (DFI)

The Dynamic Fit Index provides simulation-based cutoff values tailored to a specific model and sample size, as an alternative to generic rules of thumb for fit indices.

::: {.callout-warning}
The DFI computation is commented out because it is very time-consuming due to extensive simulations. The code below shows how to compute it if needed. These cutoffs would be valid for all subsequent analyses under the assumption that *popModel* represents the correct population model.
:::

```{r}
#| label: compute-dfi
#| eval: false
# Patient! This takes a long time due to simulations
# dynamic::catHB(pop_model, manual = TRUE, n = 1047, plot = TRUE)
```

We could calculate a DFI for each of the estimated models, assuming that the model in question would be the correct population model. However, this would lose comparability, as DFI methods for bifactor models are not yet available, and DFI for hierarchical models are still in development.

## 4-Factor CFA Model {#sec-4f}

The main model specifies four correlated latent factors (Psychological, Physical, Social, and Environment) with all items treated as ordinal. Based on preliminary analysis and theoretical considerations, items Q4 and Q3 are allowed to correlate.

### Model Specification

```{r}
#| label: specify-model-4f
# Endogenous (latent) variables are correlated by default in lavaan
# Items Q3, Q4, and Q26 have already been reverse-coded
# We allow Q4 and Q3 to have correlated residuals
spe_4fa <- "
psycho =~ Q5_P + Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
"
```

### Model Estimation

```{r}
#| label: fit-model-4f
est_4fa <- cfa(
  spe_4fa,
  data = df0,
  ordered = TRUE,      # All items are ordinal
  std.lv = TRUE,       # Standardize latent variables (variance = 1)
  estimator = "WLSMV"  # DWLS estimator for ordinal data
)
```

### Model Evaluation

::: {.callout-tip}
## Evaluating CFA Models

When evaluating CFA models, watch for:

1. **Heywood cases**: Standardized loadings > 1 or negative variances
2. **Overall fit**: SRMR < .08, RMSEA < .06, CFI/TLI > .95 (scaled versions for WLSMV)
3. **Local fit**: Standardized estimates, residual correlations, and modification indices (MI)
4. **Reliability**: Cronbach's alpha and omega > .70 for adequate internal consistency
:::

#### Overall Fit and Parameter Estimates

```{r}
#| label: summary-4f
summary(est_4fa,
  fit.measures = TRUE,   # Include fit indices
  standardized = TRUE,   # Include standardized estimates
  rsq = TRUE             # Include R-squared for items
)
```

#### Modification Indices

Modification indices (MI) suggest which parameters, if freed, would improve model fit the most. An MI > 3.84 is statistically significant at α = .05.

```{r}
#| label: tbl-mi-4f
#| tbl-cap: "Top modification indices for the 4-factor model"
knitr::kable(modificationindices(est_4fa,
  sort. = TRUE,           # Sort by MI value
  minimum.value = 3.84,   # Only show significant MIs
  maximum.number = 20     # Limit to top 20
), digits = 3)
```

::: {.callout-warning}
## Interpreting Modification Indices

High MI values often indicate local misfit. However, not all high MIs should be addressed—only those that make theoretical sense. In this model, we observe:

- High MI for correlations between physical and psychological factors
- Suggestions to allow Q5 to load on multiple factors (a red flag!)
- Q5 appears in many high MIs, indicating it may be problematic
:::

#### Residual Correlations

```{r}
#| label: residual-cor-4f
# Correlation matrix of residuals
residuals(est_4fa, type = "cor")
```

```{r}
#| label: residual-std-4f
# Standardized residuals (values > |2| indicate poor local fit)
residuals(est_4fa, type = "standardized")$cov
```

#### Reliability Coefficients

```{r}
#| label: tbl-reliability-4f
#| tbl-cap: "Reliability coefficients for the 4-factor model"
# alpha = Cronbach's alpha
# omega = Composite reliability (preferred for CFA)
# omega2 = Bentler's omega
# omega3 = McDonald's omega  
# avevar = Average Variance Extracted (should be > .5)
knitr::kable(semTools::reliability(est_4fa), digits = 3)
```

::: {.callout-note}
All reliability coefficients exceed .70, indicating adequate internal consistency for all four factors. The AVE (average variance extracted) values are also acceptable, with most factors explaining substantial variance in their indicators.
:::

#### Preliminary Evaluation

Based on the overall and local fit indices:

- **Overall fit is mediocre**, compatible with a Level 2 misspecification according to DFI standards
- **Local indices indicate**:
  - High correlation between physical and psychological domains (which may suggest a hierarchical structure)
  - Low factor loadings for Q4 (Physical) and Q24 (Environment)
  - **Warning**: Item Q5 (Psychological) appears problematic—it shows up in many high MIs and may not fit well in this factor structure

### Path Diagrams

Path diagrams provide a visual representation of the CFA model, showing factor loadings and correlations.

#### Tree Layout

```{r}
#| label: fig-diagram-4f-tree
#| fig-cap: "Path diagram of the 4-factor CFA model (tree layout)"
#| fig-width: 7
#| fig-height: 8
# Convert to semPlotModel first to avoid parsing errors with large ordinal models
sem_model_4fa <- semPlot::semPlotModel(est_4fa)

# Create path diagram with standardized estimates
semPlot::semPaths(
  sem_model_4fa,
  what = "std",              # Show standardized estimates
  style = "lisrel",          # LISREL-style layout
  layout = "tree",           # Tree layout
  intercepts = FALSE,        # Don't show intercepts
  thresholds = FALSE,        # Don't show thresholds (too cluttered)
  residuals = FALSE,         # Don't show residual variances
  edge.label.cex = 0.6,      # Edge label size
  label.cex = 1.3,           # Node label size
  weighted = FALSE,          # Don't weight edges by magnitude
  edge.color = "black",      # Black edges
  node.width = 0.6,          # Node width
  mar = c(3, 1, 3, 1)        # Margins
)
```

#### Circle Layout

```{r}
#| label: fig-diagram-4f-circle
#| fig-cap: "Path diagram of the 4-factor CFA model (circle layout)"
#| fig-width: 7
#| fig-height: 7
semPlot::semPaths(
  sem_model_4fa,
  what = "std",
  style = "lisrel",
  layout = "circle",         # Circle layout shows correlations well
  intercepts = FALSE,
  thresholds = FALSE,
  residuals = FALSE,
  edge.label.cex = 1,
  label.cex = 1,
  weighted = FALSE,
  edge.color = "black",
  mar = c(1.5, 0.8, 3, 0.3)
)
```

## 4-Factor Bifactor Model {#sec-bifactor}

A bifactor model adds a general Quality of Life (QOL) factor loading on all items, alongside the four specific group factors (Psychological, Physical, Social, Environment). In bifactor models, all factors are orthogonal (uncorrelated).

::: {.callout-note}
## Why Bifactor Models?

Bifactor models are useful when you hypothesize both a general construct (overall QOL) and specific domain constructs. However, they come with important assumptions and limitations (discussed later).
:::

Item Q5 is excluded from this model due to convergence problems—another indication that this item is problematic.

### Model Specification

```{r}
#| label: specify-model-bifactor
# In bifactor models:
# - All factors are orthogonal (uncorrelated)
# - The general factor (QOL) must be specified FIRST
# - Q4 and Q3 correlation is removed (both load on the general factor)
spe_4f_bi <- "
QOL =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P + Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F + Q20_S + Q21_S + Q22_S + Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
"
```

### Model Estimation

```{r}
#| label: fit-model-bifactor
est_4f_bi <- cfa(spe_4f_bi,
  data = df0,
  orthogonal = TRUE,   # All factors are uncorrelated (required for bifactor)
  ordered = TRUE,      # All items are ordinal
  std.lv = TRUE,       # Standardize latent variables
  estimator = "WLSMV"  # DWLS estimator for ordinal data
)
```

### Model Evaluation

#### Overall Fit and Parameter Estimates

```{r}
#| label: summary-bifactor
summary(est_4f_bi, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

#### Modification Indices

```{r}
#| label: tbl-mi-bifactor
#| tbl-cap: "Top modification indices for the bifactor model"
knitr::kable(modificationindices(est_4f_bi,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
), digits = 3)
```

#### Residual Correlations

```{r}
#| label: residual-cor-bifactor
residuals(est_4f_bi, type = "cor")
```

```{r}
#| label: residual-std-bifactor
residuals(est_4f_bi, type = "standardized")$cov
```

#### Reliability Coefficients

For bifactor models, omega coefficients from `psych::omegaFromSem()` provide the most appropriate reliability estimates, particularly omega hierarchical (ωH), which estimates the reliability attributable to the general factor alone.

```{r}
#| label: reliability-bifactor
# Omega coefficients for bifactor model
psych::omegaFromSem(fit = est_4f_bi)
```

#### Preliminary Evaluation

- **Overall fit has improved** compared to the original 4-factor model, but remains mediocre (Level 2 of DFI)
- The **general (G) factor is statistically relevant** and explains substantial variance in the observed variables
- The **Psychological specific factor loses importance** when the general factor is included—most of its variance is captured by the general QOL factor
- This pattern suggests that quality of life may be better conceptualized as a hierarchical construct

### Path Diagram

```{r}
#| label: fig-diagram-bifactor
#| fig-cap: "Path diagram of the bifactor CFA model"
#| fig-width: 8
#| fig-height: 10
# Convert to semPlotModel first
sem_model_bi <- semPlot::semPlotModel(est_4f_bi)

# Bifactor diagrams use tree3 layout for best visualization
semPlot::semPaths(
  sem_model_bi,
  what = "std",
  style = "lisrel",
  layout = "tree3",          # tree3 layout is best for bifactor models
  intercepts = FALSE,
  thresholds = FALSE,
  residuals = FALSE,
  edge.label.cex = 1.1,
  label.cex = 1.2,
  weighted = FALSE,
  edge.color = "black",
  bifactor = "QOL",          # Specify the general factor
  edge.label.position = 0.7,
  mar = c(3, 0.8, 3, 1)
)
```

## Second-Order CFA Model {#sec-2nd}

A second-order (hierarchical) model specifies that a general QOL factor loads on the four first-order factors (Psychological, Physical, Social, Environment), which in turn load on the observed items. This differs from the bifactor model where the general factor directly loads on the items.

Item Q5 is excluded for comparability with the bifactor model. The psychological factor variance must be fixed to zero due to convergence problems (empirical underidentification).

### Model Specification

```{r}
#| label: specify-model-2or
# Second-order model:
# - First-order factors load on items
# - General QOL factor loads on first-order factors
# - First-order factors can be correlated (but this is absorbed in the second-order structure)
spe_4f_2or <- "
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
QOL =~ psycho + physical + social + environment
Q3_F ~~ Q4_F
psycho ~~ 0*psycho  # Fix psycho variance to 0 due to convergence issues
"
```

::: {.callout-warning}
## Convergence Issues

Fixing the psychological factor variance to zero is necessary for model convergence. This suggests that nearly all variance in the psychological domain is explained by the general QOL factor, consistent with our bifactor results where the psychological specific factor had minimal importance.
:::

### Model Estimation

```{r}
#| label: fit-model-2or
est_4f_2or <- cfa(
  spe_4f_2or,
  data = df0,
  ordered = TRUE,      # All items are ordinal
  estimator = "WLSMV"  # DWLS estimator for ordinal data
  # Note: std.lv not used here due to the second-order structure
)
```

### Model Evaluation

#### Overall Fit and Parameter Estimates

```{r}
#| label: summary-2or
summary(est_4f_2or, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

#### Modification Indices

```{r}
#| label: tbl-mi-2or
#| tbl-cap: "Top modification indices for the second-order model"
knitr::kable(modificationindices(est_4f_2or,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
), digits = 3)
```

#### Residual Correlations

```{r}
#| label: residual-cor-2or
residuals(est_4f_2or, type = "cor")
```

```{r}
#| label: residual-std-2or
residuals(est_4f_2or, type = "standardized")$cov
```

#### Reliability Coefficients

```{r}
#| label: tbl-reliability-2or
#| tbl-cap: "Reliability coefficients for first-order factors"
knitr::kable(semTools::reliability(est_4f_2or), digits = 3)
```

For second-order models, we also need to examine the reliability of the second-order factor:

```{r}
#| label: tbl-reliability-2or-l2
#| tbl-cap: "Second-order factor reliability"
knitr::kable(as.data.frame(t(semTools::reliabilityL2(est_4f_2or, "QOL"))), digits = 3)
```

### Path Diagram

```{r}
#| label: fig-diagram-2or
#| fig-cap: "Path diagram of the second-order CFA model"
#| fig-width: 7
#| fig-height: 9
# Convert to semPlotModel first
sem_model_2or <- semPlot::semPlotModel(est_4f_2or)

semPlot::semPaths(
  sem_model_2or,
  what = "std",
  style = "lisrel",
  layout = "tree",
  intercepts = FALSE,
  thresholds = FALSE,
  residuals = FALSE,
  edge.label.cex = 0.8,
  label.cex = 1.2,
  weighted = FALSE,
  edge.color = "black",
  node.width = 0.8,
  mar = c(4, 1, 3, 1)
)
```

## 4-Factor Model Without Item Q5 {#sec-4fb}

Based on the preliminary evaluation of the original 4-factor model, item Q5 was identified as problematic. It appeared in many high modification indices, caused convergence problems in the bifactor model, and showed inconsistent behavior across models.

This section presents the revised 4-factor model excluding Q5, which will serve as our recommended model for comparison purposes.

### Model Specification

```{r}
#| label: specify-model-4fb
# Same as original 4-factor model, but without Q5
spe_4fb <- "
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
"
```

### Model Estimation

```{r}
#| label: fit-model-4fb
est_4fb <- cfa(
  spe_4fb,
  data = df0,
  ordered = TRUE,      # All items are ordinal
  std.lv = TRUE,       # Standardize latent variables
  estimator = "WLSMV"  # DWLS estimator for ordinal data
)
```

### Model Evaluation

::: {.callout-tip}
## What to Look For

When evaluating this revised model, compare:

1. Overall fit to the original model (should improve)
2. Whether modification indices still suggest problems
3. Whether factor loadings are all acceptable (> .40)
4. Whether reliability remains adequate
:::

#### Overall Fit and Parameter Estimates

```{r}
#| label: summary-4fb
summary(est_4fb, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

#### Modification Indices

```{r}
#| label: tbl-mi-4fb
#| tbl-cap: "Top modification indices for the revised 4-factor model"
knitr::kable(modificationindices(est_4fb,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
), digits = 3)
```

#### Residual Correlations

```{r}
#| label: residual-cor-4fb
residuals(est_4fb, type = "cor")
```

```{r}
#| label: residual-std-4fb
residuals(est_4fb, type = "standardized")$cov
```

#### Reliability Coefficients

```{r}
#| label: tbl-reliability-4fb
#| tbl-cap: "Reliability coefficients for the revised 4-factor model"
knitr::kable(semTools::reliability(est_4fb), digits = 3)
```

::: {.callout-note}
Notice that reliability for the Psychological domain remains adequate even after removing Q5, confirming that Q5 was not essential for measuring this construct.
:::

### Path Diagram

```{r}
#| label: fig-diagram-4fb
#| fig-cap: "Path diagram of the revised 4-factor CFA model (without Q5)"
#| fig-width: 7
#| fig-height: 7
# Convert to semPlotModel first
sem_model_4fb <- semPlot::semPlotModel(est_4fb)

semPlot::semPaths(
  sem_model_4fb,
  what = "std",
  style = "lisrel",
  layout = "circle",
  intercepts = FALSE,
  thresholds = FALSE,
  residuals = FALSE,
  edge.label.cex = 1,
  label.cex = 1,
  weighted = FALSE,
  edge.color = "black",
  mar = c(1.5, 0.8, 3, 0.3)
)
```

## Model Comparison {#sec-comparison}

Now we compare fit indices across all four CFA models to determine which best represents the WHOQOL-BREF factor structure.

### Extract Fit Indices

```{r}
#| label: fit-indices-individual
# Extract key fit indices for each model
fit_4fa    <- fitmeasures(est_4fa,    c("srmr", "rmsea.scaled", "cfi.scaled"))
fit_4f_bi  <- fitmeasures(est_4f_bi,  c("srmr", "rmsea.scaled", "cfi.scaled"))
fit_4f_2or <- fitmeasures(est_4f_2or, c("srmr", "rmsea.scaled", "cfi.scaled"))
fit_4fb    <- fitmeasures(est_4fb,    c("srmr", "rmsea.scaled", "cfi.scaled"))

# Display each model's fit
cat("4-Factor (all items):\n")
print(fit_4fa)
cat("\nBifactor:\n")
print(fit_4f_bi)
cat("\nSecond-Order:\n")
print(fit_4f_2or)
cat("\n4-Factor (no Q5):\n")
print(fit_4fb)
```

### Comparison Table

```{r}
#| label: tbl-model-comparison
#| tbl-cap: "Fit indices comparison across all CFA models"
# Create comparison table
comparison_df <- data.frame(
  Model = c("4-Factor (all items)", "Bifactor", "Second-Order", "4-Factor (no Q5)"),
  srmr = c(fit_4fa["srmr"], fit_4f_bi["srmr"], 
           fit_4f_2or["srmr"], fit_4fb["srmr"]),
  rmsea.scaled = c(fit_4fa["rmsea.scaled"], fit_4f_bi["rmsea.scaled"],
                   fit_4f_2or["rmsea.scaled"], fit_4fb["rmsea.scaled"]),
  cfi.scaled = c(fit_4fa["cfi.scaled"], fit_4f_bi["cfi.scaled"],
                 fit_4f_2or["cfi.scaled"], fit_4fb["cfi.scaled"])
)
rownames(comparison_df) <- NULL

knitr::kable(comparison_df, digits = 3)
```

### General Considerations

The models exhibit a similar level of misspecification, although the original 4-factor model (with all items) performs slightly worse. The bifactor model shows a marginally better fit, which aligns with findings in the literature. This is likely due to the greater number of parameters in bifactor structures, which generally lead to improved model fit.

Some authors argue that bifactor models have been used to "salvage" traditional scales that are no longer sustainable under modern estimation and factor selection techniques. The improved fit comes at the cost of strong assumptions and complex interpretation.

::: {.callout-important}
## Model Selection Recommendation

A known limitation of bifactor models is the assumption that specific factors are uncorrelated—an assumption that often does not hold in practice. Additionally, bifactor models can suffer from empirical underidentification and may capitalize on chance variations in the data.

The second-order model also showed convergence problems (requiring the psychological factor variance to be fixed to zero).

**Based on empirical evidence and practical considerations, the four correlated factor model excluding item Q5 is recommended** for measuring WHOQOL-BREF in this sample. This model:

- Has acceptable fit
- Avoids strong orthogonality assumptions
- Is easier to interpret
- Has no convergence issues
- Maintains adequate reliability across all domains
:::

## Post Hoc Power Analysis {#sec-power}

Power analysis assesses whether our sample size (N = 1,047) provides adequate statistical power to detect misfit and estimate parameters precisely. We use Monte Carlo simulation via the `simsem` package, simulating data from the fitted model and re-estimating it many times.

::: {.callout-note}
## Why Post Hoc Power Analysis?

While *a priori* power analysis is ideal for study planning, *post hoc* power analysis helps us understand:

1. Whether our estimates are stable and replicable
2. Whether we have adequate power to detect misfit
3. Which parameters may need larger samples for precise estimation
:::

We use the revised four-factor model (without Q5) as our generating model. To preserve the ordinal nature of items, we generate data from the fitted lavaan object rather than a simplified population model.

### Define Cutoff Values

```{r}
#| label: define-cutoffs
# Rule-of-thumb cutoffs for fit indices
rule_of_thumb <- c(srmr = 0.06, rmsea.scaled = 0.06, cfi.scaled = 0.95)
```

### Run Simulation

::: {.callout-warning}
To reduce computational time, we run only 100 replications. For final results, increase to 1,000 or even 5,000 replications. More replications provide more stable power estimates but take much longer to run.
:::

```{r}
#| label: run-simulation
#| cache: true
my_modelci <- simsem::sim(
  100,                      # Number of replications (increase for final analysis)
  n = nrow(df0),           # Sample size (1,047)
  model = spe_4fb,         # Model specification
  generate = est_4fb,      # Generate data from fitted model (preserves ordinal structure)
  lavaanfun = "cfa",       # Use cfa() function for estimation
  ordered = TRUE,          # Items are ordinal
  std.lv = TRUE,           # Standardize latent variables
  estimator = "WLSMV",     # DWLS estimator
  seed = 123456            # For reproducibility
)
```

### Simulation Results

#### Population Model Summary

```{r}
#| label: pop-summary-pow
# Shows the "true" population values used for data generation
summaryPopulation(my_modelci)
```

#### Overall Summary

```{r}
#| label: sim-summary
# Summary of simulation results
summary(my_modelci)
```

::: {.callout-note}
The summary shows power for detecting parameter differences from zero. Parameters with power < .80 may need larger samples for reliable detection. Note that only two threshold parameters have power < 0.8, and thresholds are generally not parameters of interest in CFA.
:::

#### Short Summary

```{r}
#| label: sim-short-summary
# Condensed summary
summaryShort(my_modelci)
```

### Fit Index Cutoffs

Simulation-based cutoffs at α = .05 provide an alternative to rule-of-thumb values, tailored to this specific model and sample size.

```{r}
#| label: tbl-cutoffs
#| tbl-cap: "Fit index cutoffs at alpha = 0.05"
knitr::kable(as.data.frame(t(getCutoff(my_modelci, 0.05,
  usedFit = c("srmr", "rmsea.scaled", "cfi.scaled")))), digits = 4)
```

### Power to Detect Misfit

This table shows the proportion of replications where fit indices exceeded rule-of-thumb cutoffs, indicating that the model would be rejected as having poor fit.

```{r}
#| label: tbl-power-fit
#| tbl-cap: "Proportion of replications indicating worse fit than rule-of-thumb cutoffs"
knitr::kable(as.data.frame(t(getPowerFit(my_modelci, cutoff = rule_of_thumb,
  usedFit = c("srmr", "rmsea.scaled", "cfi.scaled")))), digits = 4)
```

::: {.callout-tip}
## Interpreting Power to Detect Misfit

Low values (near 0) indicate that the model fits well in most replications—we have low power to reject it, which is actually what we want for a well-fitting model. High values would indicate we have high power to detect that the model doesn't fit well.
:::

### Sampling Distributions of Fit Indices

Visual representation of fit index distributions helps understand the variability and precision of fit assessment.

```{r}
#| label: fig-power-fit
#| fig-cap: "Sampling distributions of fit indices — Power of rejecting misspecified models"
#| fig-width: 10
#| fig-height: 4
plotPowerFit(my_modelci, cutoff = rule_of_thumb, logistic = FALSE, 
             usedFit = c("srmr", "rmsea.scaled", "cfi.scaled"))
```

```{r}
#| label: fig-cutoff-plot
#| fig-cap: "Sampling distributions of fit indices with cutoffs at alpha = 0.05"
#| fig-width: 10
#| fig-height: 4
plotCutoff(my_modelci, 0.05, usedFit = c("srmr", "rmsea.scaled", "cfi.scaled"))
```

### Detailed Parameter Summary

This table provides comprehensive information about each parameter, including:

- **Population values** (true generating values)
- **Average estimates** across replications
- **Standard deviation** of estimates
- **Average standard error** from model output
- **Coverage** (proportion of 95% CIs containing true value, should be ~.95)
- **Power** (proportion rejecting H₀: parameter = 0)
- **Relative bias** in estimates and standard errors

```{r}
#| label: param-summary
summaryParam(my_modelci, detail = TRUE, digits = 3)
```

::: {.callout-tip}
## Interpreting Parameter Summary

Look for:

- **Power > .80**: Adequate power to detect non-zero parameters
- **Coverage 0.91–0.98**: Proper confidence interval performance  
- **Rel Bias < |0.10|**: Estimates are unbiased
- **Rel SE Bias < |0.10|**: Standard errors are accurate

All parameters of interest outside these ranges deserve attention. Thresholds are generally not parameters of interest and can be ignored for interpretation purposes.
:::

## Summary and Conclusions

This tutorial demonstrated a complete CFA workflow for ordinal data using the WHOQOL-BREF instrument:

1. **Data preparation**: Loaded data directly from Mendeley Data repository
2. **Population model**: Defined a teaching example based on @lin2022  
3. **Model specification**: Tested four competing models:
   - Original 4-factor model (all items)
   - Bifactor model with general QOL factor
   - Second-order hierarchical model
   - Revised 4-factor model (excluding problematic item Q5)
4. **Model evaluation**: Used DWLS estimation appropriate for ordinal data, examining overall fit, local fit, reliability, and path diagrams
5. **Model comparison**: Systematically compared all models on key fit indices
6. **Power analysis**: Assessed statistical power via Monte Carlo simulation

::: {.callout-important}
## Key Findings and Recommendations

1. **Item Q5 is problematic** across multiple models and should be excluded
2. **The revised 4-factor correlated model** (without Q5) is recommended as the best balance of fit, interpretability, and statistical properties
3. **All four domains** (Psychological, Physical, Social, Environment) show adequate reliability
4. **Sample size (N = 1,047)** provides adequate power for parameter estimation
5. **DWLS/WLSMV estimation** is essential for proper analysis of ordinal Likert data
:::

### Pedagogical Takeaways

This tutorial illustrated several important principles for conducting "typical CFA" in applied settings:

- Always use appropriate estimators for ordinal data (DWLS/WLSMV, not ML)
- Examine both overall and local fit—overall indices can mask specific problems
- Be cautious with modification indices; only make theoretically justified changes
- Consider multiple competing models rather than accepting the first specification
- Document and report problematic items transparently
- Use simulation-based power analysis when feasible
- Bifactor models often fit better but require strong assumptions and careful interpretation

For further methodological guidance, see @rogers2022 and @rogers2024.

## References {.unnumbered}

::: {#refs}
:::
