---
title: "Setup and Data Loading"
description: "Load packages, configure environment, and import the WHOQOL-BREF dataset"
format:
  html:
    code-fold: show
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This notebook prepares the R environment and loads the WHOQOL-BREF dataset
used throughout the CFA tutorial. The data come from @rogers2021, available
on Mendeley Data, and support the reference article @rogers2022.

## Loading Packages

```{r}
#| label: load-packages
library(lavaan)
library(semTools)
library(semPlot)
library(psych)
library(simsem)
library(MASS)
library(dynamic)
```

## Environment Preparation

```{r}
#| label: set-options
options(max.print = 1000000)
set.seed(123456)
```

## Source Utility Functions

```{r}
#| label: source-utils
source("../R/formatting-functions.R")
source("../R/plot-themes.R")
```

## Step 1: ETL — Extract, Transform, Load

See reference article [@rogers2022] and database [@rogers2021].

### Reading the Database from the Internet

```{r}
#| label: load-data
# Defining the database URL
url_data <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/b58a7054-4978-4239-81db-00f9efd86e44/WHOQOL_Data.dat"

# Reading data without variable names
df0 <- read.table(url_data)

# Defining the URL of variable names
url_labels <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/ae2a5c01-2fc2-43e5-b31c-8db8cc5bff8a/WHOQOL_Labels.txt"

# Reading variable names in the first column
column_labels <- read.table(url_labels)
column_labels <- column_labels$V1

# Assigning variable names
names(df0) <- column_labels
```

### Viewing the Database

```{r}
#| label: tbl-data-head
#| tbl-cap: "First rows of the WHOQOL-BREF dataset"
head(df0)
```

```{r}
#| label: view-variable-names
names(df0)
```

```{r}
#| label: view-data-structure
str(df0)
```

```{r}
#| label: tbl-descriptives
#| tbl-cap: "Descriptive statistics for all WHOQOL-BREF items"
describe(df0)
```

### Save Data for Other Notebooks

```{r}
#| label: save-data
saveRDS(df0, "../data/whoqol_data.rds")
```

::: {.callout-note collapse="true"}
## Session Info

```{r}
#| label: session-info-setup
sessionInfo()
```
:::

## References {.unnumbered}

::: {#refs}
:::
