---
title: "Population Model and Dynamic Fit Index"
description: "Define the population model from Lin & Yao (2022) and compute dynamic fit cutoffs"
format:
  html:
    code-fold: show
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

The Dynamic Fit Index (DFI) provides simulation-based cutoff values tailored to
a specific model and sample size, as an alternative to generic rules of thumb
for fit indices.

For the calculation of the DFI and the *a priori* power analysis, the model
resulting from the meta-analysis by @lin2022, referred to as *popModel*, is
considered. In their study, @lin2022 validated the factor structure of the
WHOQOL-BREF using a meta-analysis of exploratory factor analyses combined with
social network analysis.

## Setup

```{r}
#| label: load-packages-pop
library(lavaan)
library(dynamic)
```

```{r}
#| label: source-utils-pop
source("../R/formatting-functions.R")
```

## Population Model

Given the ordinal nature of the variables, equidistant thresholds are assumed,
following a linearity assumption, with proportions of 12%, 23%, 31%, and 12%. It
is unlikely that thresholds will be available in previous studies; however, if
response frequencies for each category of each item are provided, thresholds can
be estimated using the inverse normal distribution. For a five‑point Likert
scale (1 to 5), typical equidistant thresholds would be −1.2, −0.4, 0.4, and
1.2.

Using the factor loadings obtained by @lin2022 can be questioned, since most
studies included in their meta-analysis relied on the widely criticized default
combination of principal components, Kaiser criterion, and varimax rotation, as
discussed by @rogers2022. Moreover, the evidence comes from exploratory factor
analysis rather than confirmatory factor analysis.

::: {.callout-important}
The example model (*popModel*) should be understood as a teaching illustration.
Ideally, one should look for a robust national survey that has used the WHOQOL
with CFA for ordinal data (or otherwise), providing more appropriate starting
values.
:::

```{r}
#| label: define-pop-model
pop_model <- "
 # define factors and set loadings according to Lin and Yao (2022)
psycho =~ 0.92*Q5_P + 0.81*Q6_P + 0.94*Q7_P + 0.73*Q11_P + 0.75*Q19_P + 0.63*Q26_P
physical =~ 0.84*Q3_F + 0.83*Q4_F + 0.63*Q10_F + 0.71*Q15_F + 0.68*Q16_F
            + 0.89*Q17_F + 0.61*Q18_F
social =~ 0.94*Q20_S + 0.89*Q21_S + 0.78*Q22_S
environment =~ 0.37*Q8_A + 0.49*Q9_A + 0.79*Q12_A + 0.81*Q13_A + 0.77*Q14_A
               + 0.75*Q23_A + 0.64*Q24_A + 0.73*Q25_A
 # set the maximum acceptable correlation between Q4 and Q3
Q4_F ~~ 0.3*Q3_F
 # set factor variances to 1 and correlations to 0.3
 # according to Lin and Yao (2022), correlations between factors ranged from 0.08 to 0.3
psycho ~~ 1*psycho + 0.3*physical + 0.3*social + 0.3*environment
physical ~~ 1*physical + 0.3*social + 0.3*environment
social ~~ 1*social + 0.3*environment
environment ~~ 1*environment
 # set all item residual variances according defined loadings (1-L^2)
 # except Q3 and Q4 due to the predicted correlation between the two
Q5_P ~~ 0.15*Q5_P
Q6_P ~~ 0.34*Q6_P
Q7_P ~~ 0.12*Q7_P
Q11_P ~~ 0.47*Q11_P
Q19_P ~~ 0.44*Q19_P
Q26_P ~~ 0.6*Q26_P
Q10_F ~~ 0.6*Q10_F
Q15_F ~~ 0.5*Q15_F
Q16_F ~~ 0.54*Q16_F
Q17_F ~~ 0.21*Q17_F
Q18_F ~~ 0.63*Q18_F
Q20_S ~~ 0.12*Q20_S
Q21_S ~~ 0.21*Q21_S
Q22_S ~~ 0.39*Q22_S
Q8_A ~~ 0.86*Q8_A
Q9_A ~~ 0.76*Q9_A
Q12_A ~~ 0.38*Q12_A
Q13_A ~~ 0.34*Q13_A
Q14_A ~~ 0.41*Q14_A
Q23_A ~~ 0.44*Q23_A
Q24_A ~~ 0.59*Q24_A
Q25_A ~~ 0.47*Q25_A
 # set thresholds
Q5_P | -1.2*t1
Q5_P | -0.4*t2
Q5_P | 0.4*t3
Q5_P | 1.2*t4
Q6_P | -1.2*t1
Q6_P | -0.4*t2
Q6_P | 0.4*t3
Q6_P | 1.2*t4
Q7_P | -1.2*t1
Q7_P | -0.4*t2
Q7_P | 0.4*t3
Q7_P | 1.2*t4
Q11_P | -1.2*t1
Q11_P | -0.4*t2
Q11_P | 0.4*t3
Q11_P | 1.2*t4
Q19_P | -1.2*t1
Q19_P | -0.4*t2
Q19_P | 0.4*t3
Q19_P | 1.2*t4
Q26_P | -1.2*t1
Q26_P | -0.4*t2
Q26_P | 0.4*t3
Q26_P | 1.2*t4
Q3_F | -1.2*t1
Q3_F | -0.4*t2
Q3_F | 0.4*t3
Q3_F | 1.2*t4
Q4_F | -1.2*t1
Q4_F | -0.4*t2
Q4_F | 0.4*t3
Q4_F | 1.2*t4
Q10_F | -1.2*t1
Q10_F | -0.4*t2
Q10_F | 0.4*t3
Q10_F | 1.2*t4
Q15_F | -1.2*t1
Q15_F | -0.4*t2
Q15_F | 0.4*t3
Q15_F | 1.2*t4
Q16_F | -1.2*t1
Q16_F | -0.4*t2
Q16_F | 0.4*t3
Q16_F | 1.2*t4
Q17_F | -1.2*t1
Q17_F | -0.4*t2
Q17_F | 0.4*t3
Q17_F | 1.2*t4
Q18_F | -1.2*t1
Q18_F | -0.4*t2
Q18_F | 0.4*t3
Q18_F | 1.2*t4
Q20_S | -1.2*t1
Q20_S | -0.4*t2
Q20_S | 0.4*t3
Q20_S | 1.2*t4
Q21_S | -1.2*t1
Q21_S | -0.4*t2
Q21_S | 0.4*t3
Q21_S | 1.2*t4
Q22_S | -1.2*t1
Q22_S | -0.4*t2
Q22_S | 0.4*t3
Q22_S | 1.2*t4
Q8_A | -1.2*t1
Q8_A | -0.4*t2
Q8_A | 0.4*t3
Q8_A | 1.2*t4
Q9_A | -1.2*t1
Q9_A | -0.4*t2
Q9_A | 0.4*t3
Q9_A | 1.2*t4
Q12_A | -1.2*t1
Q12_A | -0.4*t2
Q12_A | 0.4*t3
Q12_A | 1.2*t4
Q13_A | -1.2*t1
Q13_A | -0.4*t2
Q13_A | 0.4*t3
Q13_A | 1.2*t4
Q14_A | -1.2*t1
Q14_A | -0.4*t2
Q14_A | 0.4*t3
Q14_A | 1.2*t4
Q23_A | -1.2*t1
Q23_A | -0.4*t2
Q23_A | 0.4*t3
Q23_A | 1.2*t4
Q24_A | -1.2*t1
Q24_A | -0.4*t2
Q24_A | 0.4*t3
Q24_A | 1.2*t4
Q25_A | -1.2*t1
Q25_A | -0.4*t2
Q25_A | 0.4*t3
Q25_A | 1.2*t4
"
```

### Save Population Model

```{r}
#| label: save-pop-model
saveRDS(pop_model, "../models/pop_model.rds")
```

## Dynamic Fit Index

::: {.callout-warning}
The DFI computation is commented out because it is very time-consuming due to
simulations. Uncomment and run if you need to recompute the cutoffs.
:::

```{r}
#| label: compute-dfi
#| eval: false
# Patient! It takes a while due to simulations
# dynamic::catHB(pop_model, manual = TRUE, n = 1047, plot = TRUE)
```

These cutoffs were valid for all subsequent analyses, as we are assuming that
*popModel* is the correct population model.

We could calculate a DFI for each of the estimated models, assuming that the
model in question would be the correct population model, as is usually the case
(see the referenced McNeish and Wolf papers). But in this case, we would lose
comparability, as to date, we do not have DFI for bifactor models, and DFI for
hierarchical models are in progress.

::: {.callout-note collapse="true"}
## Session Info

```{r}
#| label: session-info-pop
sessionInfo()
```
:::

## References {.unnumbered}

::: {#refs}
:::
