---
title: "4-Factor CFA Model"
description: "Specify, estimate, and evaluate the correlated 4-factor model for WHOQOL-BREF"
format:
  html:
    code-fold: show
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This notebook fits the main correlated four-factor CFA model to the
WHOQOL-BREF data. The four factors are: Psychological, Physical, Social, and
Environment. All items are treated as ordinal and estimated with DWLS (WLSMV).

## Setup

```{r}
#| label: load-packages-4f
library(lavaan)
library(semTools)
library(semPlot)
```

```{r}
#| label: source-utils-4f
source("../R/formatting-functions.R")
source("../R/plot-themes.R")
```

```{r}
#| label: load-data-4f
df0 <- readRDS("../data/whoqol_data.rds")
```

## Step 2: Model Specification

-   Endogenous variables are correlated by default
-   The items Q3, Q4 and Q26 have already been reversed
-   With items Q4 and Q3 correlated

```{r}
#| label: specify-model-4f
spe_4fa <- "
psycho =~ Q5_P + Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
"
```

## Step 3: Model Estimation

-   With the CFA function
-   With WLSMV/DWLS estimator
-   Fixing the variance to standardize the factor scale
-   All items are ordinal

```{r}
#| label: fit-model-4f
est_4fa <- cfa(
  spe_4fa,
  data = df0,
  ordered = TRUE,
  std.lv = TRUE,
  estimator = "WLSMV"
)
```

```{r}
#| label: save-fit-4f
#| cache: false
saveRDS(est_4fa, "../models/est_4fa.rds")
```

## Step 4: Model Evaluation

::: {.callout-tip}
Be careful with Heywood cases: Std.all > 1 or variance < 0.
Look at SRMR, RMSEA and CFI to assess overall fit.
Look at standardized estimates, residual correlations, and MI to assess local fit.
:::

### Overall Fit and Parameter Estimates

```{r}
#| label: summary-4f
#| tbl-cap: "Summary of the 4-factor CFA model"
summary(est_4fa,
  fit.measures = TRUE,
  standardized = TRUE, rsq = TRUE
)
```

### Modification Indices (MI)

```{r}
#| label: tbl-mi-4f
#| tbl-cap: "Top modification indices for the 4-factor model"
knitr::kable(modificationindices(est_4fa,
  sort. = TRUE,
  minimum.value = 3.84,
  maximum.number = 20
), digits = 3)
```

### Residual Correlations

```{r}
#| label: residual-cor-4f
residuals(est_4fa, type = "cor")
```

```{r}
#| label: residual-std-4f
residuals(est_4fa, type = "standardized")$cov
```

### Reliability Coefficients

-   alpha = Cronbach's; omega = composite reliability; omega2 = Bentler's
-   omega3 = McDonald's; avevar = AVE (average variance extracted > 0.5)

::: {.callout-note}
Reliability coefficients must be > .7 for adequate internal consistency.
:::

```{r}
#| label: tbl-reliability-4f
#| tbl-cap: "Reliability coefficients for the 4-factor model"
knitr::kable(format_reliability(est_4fa))
```

### Preliminary Evaluation

-   The overall fit was mediocre, compatible with a Level 2 of misspecification
-   Local indices indicated:
    -   High correlation between physical and psychological
    -   Low factor loading of Q4 and Q24
    -   Warning: we may have problems with Q5 (it is the cause of many high MI)

## Model Diagrams

```{r}
#| label: fig-diagram-4f-tree
#| fig-cap: "Path diagram of the 4-factor CFA model (tree layout)"
plot_cfa_diagram(est_4fa, edge.label.cex = 0.6, label.cex = 1.3, node.width = 0.6)
```

```{r}
#| label: fig-diagram-4f-circle
#| fig-cap: "Path diagram of the 4-factor CFA model (circle layout)"
plot_cfa_diagram(est_4fa, layout = "circle",
                 edge.label.cex = 1, label.cex = 1,
                 mar = c(1.5, 0.8, 3, 0.3))
```

::: {.callout-note collapse="true"}
## Session Info

```{r}
#| label: session-info-4f
sessionInfo()
```
:::
