---
title: "Bifactor CFA Model"
description: "Specify, estimate, and evaluate the bifactor model for WHOQOL-BREF"
format:
  html:
    code-fold: show
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This notebook fits a bifactor model to the WHOQOL-BREF data, with a general
Quality of Life (QOL) factor and four specific group factors. Item Q5 is
excluded due to convergence problems.

## Setup

```{r}
#| label: load-packages-bi
library(lavaan)
library(semTools)
library(semPlot)
library(psych)
```

```{r}
#| label: source-utils-bi
source("../R/formatting-functions.R")
source("../R/plot-themes.R")
```

```{r}
#| label: load-data-bi
df0 <- readRDS("../data/whoqol_data.rds")
```

## Step 2: Model Specification

-   Endogenous variables are not correlated (orthogonal)
-   Items Q3, Q4 and Q26 have already been reversed
-   Without considering the correlated items Q4 and Q3 (the G factor affects
    both items)
-   Without item Q5 due to convergence problems
-   The G factor (QOL) must be first in the model specification

```{r}
#| label: specify-model-bifactor
spe_4f_bi <- "
QOL =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P + Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F + Q20_S + Q21_S + Q22_S + Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
"
```

## Step 3: Model Estimation

-   With the CFA function
-   With WLSMV/DWLS estimator
-   Fixing the variance to standardize the factor scale
-   All items are ordinal
-   Orthogonal factors (required for bifactor)

```{r}
#| label: fit-model-bifactor
est_4f_bi <- cfa(spe_4f_bi,
  data = df0,
  orthogonal = TRUE,
  ordered = TRUE,
  std.lv = TRUE,
  estimator = "WLSMV"
)
```

```{r}
#| label: save-fit-bifactor
saveRDS(est_4f_bi, "../models/est_4f_bi.rds")
```

## Step 4: Model Evaluation

::: {.callout-tip}
Be careful with Heywood cases: Std.all > 1 or variance < 0.
Look at SRMR, RMSEA and CFI to assess overall fit.
Look at standardized estimates, residual correlations, and MI to assess local fit.
:::

### Overall Fit and Parameter Estimates

```{r}
#| label: tbl-summary-bifactor
#| tbl-cap: "Summary of the bifactor CFA model"
summary(est_4f_bi, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

### Modification Indices (MI)

```{r}
#| label: tbl-mi-bifactor
#| tbl-cap: "Top modification indices for the bifactor model"
modificationindices(est_4f_bi,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
)
```

### Residual Correlations

```{r}
#| label: residual-cor-bifactor
residuals(est_4f_bi, type = "cor")
```

```{r}
#| label: residual-std-bifactor
residuals(est_4f_bi, type = "standardized")$cov
```

### Reliability Coefficients

For the G factor in a bifactor model, omega from `psych::omegaFromSem()` is
the appropriate reliability measure.

```{r}
#| label: tbl-reliability-bifactor
#| tbl-cap: "Reliability coefficients for the bifactor model (G factor)"
psych::omegaFromSem(fit = est_4f_bi)
```

### Preliminary Evaluation

-   Overall Fit has improved, but also mediocre (Level 2 of DFI)
-   The G factor is statistically relevant and explains a lot of the observed
    variables
-   The psycho factor loses a lot of importance

## Model Diagram

```{r}
#| label: fig-diagram-bifactor
#| fig-cap: "Path diagram of the bifactor CFA model"
plot_cfa_diagram(est_4f_bi,
                 layout = "tree3",
                 bifactor = "QOL",
                 edge.label.cex = 1.1,
                 label.cex = 1.2,
                 mar = c(3, 0.8, 3, 1))
```

::: {.callout-note collapse="true"}
## Session Info

```{r}
#| label: session-info-bifactor
sessionInfo()
```
:::
