---
title: "Second-Order CFA Model"
description: "Specify, estimate, and evaluate the second-order model for WHOQOL-BREF"
format:
  html:
    code-fold: show
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This notebook fits a second-order (hierarchical) CFA model where a general
Quality of Life factor (QOL) loads on the four first-order factors. Item Q5 is
excluded for comparability with the bifactor model.

## Setup

```{r}
#| label: load-packages-2or
library(lavaan)
library(semTools)
library(semPlot)
```

```{r}
#| label: source-utils-2or
source("../R/formatting-functions.R")
source("../R/plot-themes.R")
```

```{r}
#| label: load-data-2or
df0 <- readRDS("../data/whoqol_data.rds")
```

## Step 2: Model Specification

-   There is a suggestive correlation between endogenous variables that may
    indicate a second order model, mainly between psycho and physical
-   Items Q3, Q4 and Q26 have already been reversed
-   With items Q4 and Q3 correlated
-   Without item Q5 for comparability
-   Fixing psycho variance due to convergence problems

```{r}
#| label: specify-model-2or
spe_4f_2or <- "
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
QOL =~ psycho + physical + social + environment
Q3_F ~~ Q4_F
psycho ~~ 0*psycho
"
```

## Step 3: Model Estimation

-   With CFA function
-   With WLSMV/DWLS estimator
-   All items are ordinal

```{r}
#| label: fit-model-2or
est_4f_2or <- cfa(
  spe_4f_2or,
  data = df0,
  ordered = TRUE,
  estimator = "WLSMV"
)
```

```{r}
#| label: save-fit-2or
#| cache: false
saveRDS(est_4f_2or, "../models/est_4f_2or.rds")
```

## Step 4: Model Evaluation

::: {.callout-tip}
Be careful with Heywood cases: Std.all > 1 or variance < 0.
Look at SRMR, RMSEA and CFI to assess overall fit.
Look at standardized estimates, residual correlations, and MI to assess local fit.
:::

### Overall Fit and Parameter Estimates

```{r}
#| label: summary-2or
#| tbl-cap: "Summary of the second-order CFA model"
summary(est_4f_2or, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

### Modification Indices (MI)

```{r}
#| label: tbl-mi-2or
#| tbl-cap: "Top modification indices for the second-order model"
knitr::kable(modificationindices(est_4f_2or,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
), digits = 3)
```

### Residual Correlations

```{r}
#| label: residual-cor-2or
residuals(est_4f_2or, type = "cor")
```

```{r}
#| label: residual-std-2or
residuals(est_4f_2or, type = "standardized")$cov
```

### Reliability Coefficients

-   alpha = Cronbach's; omega = composite reliability; omega2 = Bentler's
-   omega3 = McDonald's; avevar = AVE (average variance extracted > .5)

```{r}
#| label: tbl-reliability-2or
#| tbl-cap: "Reliability coefficients for the second-order model"
knitr::kable(format_reliability(est_4f_2or))
```

```{r}
#| label: tbl-reliability-2or-l2
#| tbl-cap: "Second-order factor reliability"
knitr::kable(as.data.frame(t(semTools::reliabilityL2(est_4f_2or, "QOL"))), digits = 3)
```

## Model Diagram

```{r}
#| label: fig-diagram-2or
#| fig-cap: "Path diagram of the second-order CFA model"
plot_cfa_diagram(est_4f_2or,
                 edge.label.cex = 0.8,
                 label.cex = 1.2,
                 node.width = 0.8,
                 mar = c(4, 1, 3, 1))
```

::: {.callout-note collapse="true"}
## Session Info

```{r}
#| label: session-info-2or
sessionInfo()
```
:::
