---
title: "4-Factor Model Without Item Q5"
description: "Revised 4-factor model excluding the problematic Q5 item"
format:
  html:
    code-fold: show
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This notebook fits the revised four-factor CFA model, excluding item Q5 which
was identified as problematic in the initial model evaluation. This model is
used for comparison purposes.

## Setup

```{r}
#| label: load-packages-4fb
library(lavaan)
library(semTools)
library(semPlot)
```

```{r}
#| label: source-utils-4fb
source("../R/formatting-functions.R")
source("../R/plot-themes.R")
```

```{r}
#| label: load-data-4fb
df0 <- readRDS("../data/whoqol_data.rds")
```

## Step 2: Model Specification

-   Endogenous variables are correlated by default
-   The items Q3, Q4 and Q26 have already been reversed
-   With items Q4 and Q3 correlated
-   Without item Q5 for comparison

```{r}
#| label: specify-model-4fb
spe_4fb <- "
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
"
```

## Step 3: Model Estimation

-   With the CFA function
-   With WLSMV/DWLS estimator
-   Fixing the variance to standardize the factor scale
-   All items are ordinal

```{r}
#| label: fit-model-4fb
est_4fb <- cfa(
  spe_4fb,
  data = df0,
  ordered = TRUE,
  std.lv = TRUE,
  estimator = "WLSMV"
)
```

```{r}
#| label: save-fit-4fb
#| cache: false
saveRDS(est_4fb, "../models/est_4fb.rds")
```

## Step 4: Model Evaluation

::: {.callout-tip}
Be careful with Heywood cases: Std.all > 1 or variance < 0.
Look at the χ²/df, CFI, TLI, SRMR and RMSEA to assess overall fit.
Look at standardized estimates, residual correlations, and MI to assess local fit.
:::

### Overall Fit and Parameter Estimates

```{r}
#| label: summary-4fb
#| tbl-cap: "Summary of the revised 4-factor CFA model (without Q5)"
summary(est_4fb, fit.measures = TRUE, standardized = TRUE, rsq = TRUE)
```

### Modification Indices (MI)

```{r}
#| label: tbl-mi-4fb
#| tbl-cap: "Top modification indices for the revised 4-factor model"
knitr::kable(modificationindices(est_4fb,
  sort. = TRUE, minimum.value = 3.84,
  maximum.number = 20
), digits = 3)
```

### Residual Correlations

```{r}
#| label: residual-cor-4fb
residuals(est_4fb, type = "cor")
```

```{r}
#| label: residual-std-4fb
residuals(est_4fb, type = "standardized")$cov
```

### Reliability Coefficients

-   alpha = Cronbach's; omega = composite reliability; omega2 = Bentler's
-   omega3 = McDonald's; avevar = AVE (average variance extracted)

```{r}
#| label: tbl-reliability-4fb
#| tbl-cap: "Reliability coefficients for the revised 4-factor model"
knitr::kable(format_reliability(est_4fb))
```

## Model Diagram

```{r}
#| label: fig-diagram-4fb
#| fig-cap: "Path diagram of the revised 4-factor CFA model (without Q5)"
plot_cfa_diagram(est_4fb, layout = "circle",
                 edge.label.cex = 1, label.cex = 1,
                 mar = c(1.5, 0.8, 3, 0.3))
```

::: {.callout-note collapse="true"}
## Session Info

```{r}
#| label: session-info-4fb
sessionInfo()
```
:::
