---
title: "Model Comparison"
description: "Compare fit indices across all CFA models"
format:
  html:
    code-fold: show
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This notebook compares the fit indices of all CFA models fitted in the previous
notebooks. The comparison helps determine which model best represents the
WHOQOL-BREF factor structure.

## Setup

```{r}
#| label: load-packages-comp
library(lavaan)
```

```{r}
#| label: source-utils-comp
source("../R/formatting-functions.R")
```

## Load Fitted Models

```{r}
#| label: load-models-comp
est_4fa    <- readRDS("../models/est_4fa.rds")
est_4f_bi  <- readRDS("../models/est_4f_bi.rds")
est_4f_2or <- readRDS("../models/est_4f_2or.rds")
est_4fb    <- readRDS("../models/est_4fb.rds")
```

## Fit Indices by Model

```{r}
#| label: fit-indices-4f
fit_4fa <- fitmeasures(object = est_4fa, c("srmr", "rmsea.scaled", "cfi.scaled"))
fit_4fa
```

```{r}
#| label: fit-indices-bifactor
fit_4f_bi <- fitmeasures(object = est_4f_bi, c("srmr", "rmsea.scaled", "cfi.scaled"))
fit_4f_bi
```

```{r}
#| label: fit-indices-2or
fit_4f_2or <- fitmeasures(object = est_4f_2or, c("srmr", "rmsea.scaled", "cfi.scaled"))
fit_4f_2or
```

```{r}
#| label: fit-indices-4fb
fit_4fb <- fitmeasures(object = est_4fb, c("srmr", "rmsea.scaled", "cfi.scaled"))
fit_4fb
```

## Comparison Table

```{r}
#| label: tbl-model-comparison
#| tbl-cap: "Fit indices comparison across all CFA models"
knitr::kable(compare_fit(
  `4-Factor (all items)` = est_4fa,
  `Bifactor`             = est_4f_bi,
  `Second-Order`         = est_4f_2or,
  `4-Factor (no Q5)`     = est_4fb
))
```

## General Considerations

The models exhibit a similar level of misspecification, although the original
4-factor model (with all items) performs slightly worse. The bifactor model
shows a marginally better fit, which aligns with findings in the literature.
This is likely due to the greater number of parameters in bifactor structures,
which generally lead to improved model fit. Some authors even argue that
bifactor models have been used to salvage traditional scales that are no longer
sustainable under modern estimation and factor selection techniques.

::: {.callout-important}
A known limitation of bifactor models is the assumption that specific factors
are uncorrelated — an assumption that often does not hold in practice.
Therefore, based on empirical evidence, the four correlated factor model
excluding item Q5 is recommended.
:::

::: {.callout-note collapse="true"}
## Session Info

```{r}
#| label: session-info-comp
sessionInfo()
```
:::
