---
title: "Post Hoc Power Analysis"
description: "Post hoc power analysis via simulation for the revised 4-factor model"
format:
  html:
    code-fold: show
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This notebook performs a post hoc power analysis for the revised four-factor CFA
model (without item Q5). We use `simsem` to simulate data from the fitted model
and assess whether the sample size (N = 1,047) provides adequate statistical
power.

## Setup

```{r}
#| label: load-packages-pow
library(lavaan)
library(simsem)
```

```{r}
#| label: source-utils-pow
source("../R/formatting-functions.R")
```

```{r}
#| label: load-data-pow
df0 <- readRDS("../data/whoqol_data.rds")
```

## Load Fitted Model

```{r}
#| label: load-model-pow
est_4fb <- readRDS("../models/est_4fb.rds")
```

## Model Specification for Simulation

```{r}
#| label: specify-model-pow
spe_4fb <- "
psycho =~ Q6_P + Q7_P + Q11_P + Q19_P + Q26_P
physical =~ Q3_F + Q4_F + Q10_F + Q15_F + Q16_F + Q17_F + Q18_F
social =~ Q20_S + Q21_S + Q22_S
environment =~ Q8_A + Q9_A + Q12_A + Q13_A + Q14_A + Q23_A + Q24_A + Q25_A
Q4_F ~~ Q3_F
"
```

## Define Fit Cutoffs

```{r}
#| label: define-cutoffs
rule_of_thumb <- c(srmr = 0.06, rmsea.scaled = 0.06, cfi.scaled = 0.95)
```

## Run Simulation

::: {.callout-warning}
To reduce computational time, we run only 100 replications. For final results,
increase to 1,000 or even 5,000 replications.
:::

```{r}
#| label: run-simulation
#| cache: true
my_modelci <- simsem::sim(100,
  n = nrow(df0),
  model = spe_4fb,
  generate = est_4fb,
  lavaanfun = "cfa",
  ordered = TRUE,
  std.lv = TRUE,
  estimator = "WLSMV",
  seed = 123456
)
```

```{r}
#| label: save-simulation
#| cache: false
saveRDS(my_modelci, "../models/my_modelci.rds")
```

## Simulation Results

### Population Model Summary

```{r}
#| label: pop-summary-pow
summaryPopulation(my_modelci)
```

### Overall Summary

```{r}
#| label: sim-summary
#| tbl-cap: "Simulation summary for the power analysis"
summary(my_modelci)
```

Only two thresholds power < 0.8 (thresholds are generally not parameters of
interest).

### Short Summary

```{r}
#| label: sim-short-summary
summaryShort(my_modelci)
```

### Fit Index Cutoffs

```{r}
#| label: tbl-cutoffs
#| tbl-cap: "Fit index cutoffs at alpha = 0.05"
knitr::kable(as.data.frame(t(getCutoff(my_modelci, 0.05,
  usedFit = c("srmr", "rmsea.scaled", "cfi.scaled")))), digits = 4)
```

### Power to Detect Misfit

```{r}
#| label: tbl-power-fit
#| tbl-cap: "Proportion of replications indicating worse fit than rule-of-thumb cutoffs"
knitr::kable(as.data.frame(t(getPowerFit(my_modelci, cutoff = rule_of_thumb,
  usedFit = c("srmr", "rmsea.scaled", "cfi.scaled")))), digits = 4)
```

### Sampling Distributions of Fit Indices

```{r}
#| label: fig-power-fit
#| fig-cap: "Sampling distributions of fit indices — Power of rejecting misspecified models"
plotPowerFit(my_modelci, cutoff = rule_of_thumb, logistic = FALSE, usedFit = c("srmr", "rmsea.scaled", "cfi.scaled"))
```

```{r}
#| label: fig-cutoff-plot
#| fig-cap: "Sampling distributions of fit indices with cutoffs at alpha = 0.05"
plotCutoff(my_modelci, 0.05, usedFit = c("srmr", "rmsea.scaled", "cfi.scaled"))
```

### Detailed Parameter Summary

```{r}
#| label: param-summary
#| tbl-cap: "Parameter estimates and power across replications (alpha = .05)"
summaryParam(my_modelci, detail = TRUE, digits = 3)
```

::: {.callout-tip}
Look for Power > .8, Coverage 0.91–0.98, Rel Bias < |0.10| and Rel SE Bias < |0.10|.
All parameters of interest outside this range are points of attention.
Thresholds are generally not parameters of interest.
:::

::: {.callout-note collapse="true"}
## Session Info

```{r}
#| label: session-info-pow
sessionInfo()
```
:::
